---
title: "Normalizing Rates"
author: "HM Putnam"
date: "2023-11-26"
output: html_document
---

## Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("devtools" %in% rownames(installed.packages()) == 'FALSE') install.packages('devtools') 
library(devtools)
if ("segmented" %in% rownames(installed.packages()) == 'FALSE') install.packages('segmented') 
if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("LoLinR" %in% rownames(installed.packages()) == 'FALSE') install_github('colin-olito/LoLinR') 
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if ("chron" %in% rownames(installed.packages()) == 'FALSE') install.packages('chron') 
if ("plyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('plyr') 
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr') 
if ("stringr" %in% rownames(installed.packages()) == 'FALSE') install.packages('stringr') 
if ("Rmisc" %in% rownames(installed.packages()) == 'FALSE') install.packages('Rmisc') 
if ("respR" %in% rownames(installed.packages()) == 'FALSE') install.packages('respR') 
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("ggpubr" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggpubr') 

#load packages
library("ggplot2")
library("segmented")
library("plotrix")
library("gridExtra")
library("LoLinR")
library("lubridate")
library("chron")
library('plyr')
library('dplyr')
library('stringr')
library('Rmisc')
library('respR')
library('tidyverse')
library('emmeans')
library('ggpubr')

```

Load Raw extracted Oxygen Flux Rates
```{r, warning=FALSE, message=FALSE}
Photo.Rates <- read.csv(file = "output/20231213_pr_sdr_rates_raw.csv") #read file back in so slopes don't have to be generated every time
Photo.Rates = subset(Photo.Rates, select = -c(X) ) #remove empty column

#format  column
Photo.Rates<-Photo.Rates %>%
  mutate(Plate = str_sub(Plate, 1, -1))
Photo.Rates$Plate<-as.integer(Photo.Rates$Plate) #format as # rather than run #, set as integer
```

## Standardize and normalize  

Merge rates files with sample info for normalization, plotting, and stats.  

```{r, warning=FALSE, message=FALSE}
Sample.Info <- read_csv(file = "data/larval/pr_sdr/PR_SDR_Sample_Info.csv")%>%
  select(!Notes) %>%
  filter(Cohort=="20231212") %>%
  filter(Date=="20231213")

#Sample.Info <- Sample.Info %>%
#  filter(Date == 20231216)

#make unique ID column
Sample.Info$unique <- paste0(Sample.Info$Date, Sample.Info$Plate, Sample.Info$Chamber.ID)
str(Sample.Info)

Photo.Rates$unique <- paste0(Photo.Rates$Date, Photo.Rates$Plate, Photo.Rates$Chamber.ID)
str(Photo.Rates)

oxygen<-left_join(Photo.Rates, Sample.Info,  by="unique") #add photosynthesis data

oxygen<- Photo.Rates %>%
  left_join(Sample.Info, by="unique", suffix=c("",".y")) %>%
  select(-ends_with(".y"))

colnames(oxygen)[colnames(oxygen) == 'Intercept'] <- 'Photo.Intercept' #rename to specify R
colnames(oxygen)[colnames(oxygen) == 'umol.L.min'] <- 'P.umol.L.min' #rename to specify R

oxygen <- oxygen%>% 
  filter(!Phenotype == "No Sensor")


```

Check for Outliers in the raw data
```{r}
oxygen %>% 
  filter(Type=="Blank") %>%
  ggplot(aes(Phenotype, P.umol.L.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

oxygen %>% 
  filter(Type=="Sample") %>%
  ggplot(aes(Phenotype, P.umol.L.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

oxygen %>% 
  ggplot(aes(Type, P.umol.L.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light, Phenotype), ncol=4)+
  theme_bw()


```

Account for volume to obtain umol per minute.   
```{r, results=TRUE, warning=FALSE, message=FALSE}
#Account for chamber volume to convert from umol L-1 m-1 to umol m-1. This removes per Liter
oxygen$P.umol.min <- oxygen$P.umol.L.min * oxygen$Volume #calculate

```


Analyze the data by Cohorts without the data for testing by number of larvae 
```{r}

blank_data <- subset(oxygen, Type == "Blank") #subset to blank data only

blank_data %>% 
  ggplot(aes(Cohort, P.umol.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

#remove positive respiration blanks 
blank_data <-blank_data %>% 
  filter(!unique =="202312137B2")

blank_data %>% 
  ggplot(aes(Light, P.umol.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  theme_bw()

#display mean blank values
mean.blank.data <- blank_data %>% group_by(Plate, Light) %>%
summarise(mean = mean(P.umol.min, na.rm=TRUE)) #mean P phase blanks

mean.blank.data %>% 
  ggplot(aes(Plate, mean))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()


photo.blnk <- aggregate(P.umol.min ~ Date*Plate*Light, data=blank_data, mean) #calculate average blank during light for each run
colnames(photo.blnk)[colnames(photo.blnk) == 'P.umol.min'] <- 'P.Blank.umol.min' #rename to specify blank for R

oxygen <- full_join(oxygen, photo.blnk) #add R blanks to master
```

Subtract blank values to generate a "corrected" value for umol O2 min-1.  

```{r, warning=FALSE, message=FALSE}
oxygen$P.umol.min.corr<-oxygen$P.umol.min-oxygen$P.Blank.umol.min #subtract R blanks
```

Normalize to biologically relevant measure. Here, normalize to number of larvae. This can be substituted or changed for larval size/volume as well.  

```{r, warning=FALSE, message=FALSE}
oxygen.bio <- oxygen %>% filter(Org.Number == 60) #isolate only biological samples and 
#respiration
oxygen.bio$P.umol.org.min <- oxygen.bio$P.umol.min.corr/oxygen.bio$Org.Number #calculate oxygen per organism 
oxygen.bio$P.nmol.org.min <- oxygen.bio$P.umol.org.min*1000 #calculate nanomoles 
#oxygen.bio$P.nmol.org.hr <- oxygen.bio$P.umol.org.min*60 #calculate nanomoles per hour

```

Plot values.  
```{r}
options(scipen = 999)
oxygen.bio %>% 
  ggplot(aes(Temperature, P.umol.min.corr))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

```


Comparison of development at 28 and 31 
```{r, warning=FALSE, message=FALSE}
oxygen.bio.devo <- oxygen.bio 

devo.temp <- ggplot(oxygen.bio.devo, aes(x = Temperature, y = P.nmol.org.min)) +
  stat_summary(aes(group = Phenotype), fun.y = mean, geom = "path",  position = position_dodge(width = 0.3)) +
  stat_summary(aes(color = Phenotype), fun.data = mean_cl_boot, geom = "errorbar", width = 0.1,  position = position_dodge(width = 0.3)) +
  stat_summary(aes(color = Phenotype), fun.y = mean, geom = "point", size = 4,  position = position_dodge(width = 0.3)) +
  geom_point(aes(color = Phenotype), size = 2, position = position_dodge(width = 0.3)) +
  facet_wrap(vars(Light), ncol=2) +
  scale_color_manual(values = c("blue", "red"))+
  theme_bw()
devo.temp

oxygen.bio.devo.light <- oxygen.bio.devo %>%
  filter(Light == 1)

oxygen.bio.devo.light$Pnet <- oxygen.bio.devo.light$P.nmol.org.min

oxygen.bio.devo.dark <- oxygen.bio.devo %>%
  filter(Light == 2)

oxygen.bio.devo.dark$LEDR <- oxygen.bio.devo.dark$P.nmol.org.min

oxygen.bio.devo.wide <- left_join(oxygen.bio.devo.light, oxygen.bio.devo.dark, by="unique")

oxygen.bio.devo.wide$Pgross <- oxygen.bio.devo.wide$Pnet - oxygen.bio.devo.wide$LEDR

oxygen.bio.devo.wide$PR.ratio <- oxygen.bio.devo.wide$Pnet/oxygen.bio.devo.wide$LEDR


model.devo.Pnet <- lm(Pnet~ Temperature.x*Phenotype.x, data=oxygen.bio.devo.wide)
summary(model.devo.Pnet)

model.devo.Pgross <- lm(Pgross~ Temperature.x*Phenotype.x, data=oxygen.bio.devo.wide)
summary(model.devo.Pgross)

model.devo.dark <- lm(LEDR~ Temperature.x*Phenotype.x, data=oxygen.bio.devo.wide)
summary(model.devo.dark)

model.devo.dark <- lm(PR.ratio~ Temperature.x*Phenotype.x, data=oxygen.bio.devo.wide)
summary(model.devo.dark)
```

```{r}

Pnet <- ggplot(oxygen.bio.devo.wide, aes(x = Temperature.x, y = Pnet)) +
  stat_summary(aes(group = Phenotype.x), fun.y = mean, geom = "path",  position = position_dodge(width = 0.3)) +
  stat_summary(aes(color = Phenotype.x), fun.data = mean_cl_boot, geom = "errorbar", width = 0.1,  position = position_dodge(width = 0.3)) +
  stat_summary(aes(color = Phenotype.x), fun.y = mean, geom = "point", size = 4,  position = position_dodge(width = 0.3)) +
  geom_hline(yintercept=0,linetype=3)+
  ylim(-0.003, 0.002)+
  xlab("Temperature °C")+
  ylab("Pnet nmol/larvae/min")+
  geom_point(aes(color = Phenotype.x), size = 2, position = position_dodge(width = 0.3)) +
  scale_color_manual(values = c("blue", "red"))+
  theme_classic()+
  theme(legend.position = "none")


Pgross <- ggplot(oxygen.bio.devo.wide, aes(x = Temperature.x, y = Pgross)) +
  stat_summary(aes(group = Phenotype.x), fun.y = mean, geom = "path",  position = position_dodge(width = 0.3)) +
  stat_summary(aes(color = Phenotype.x), fun.data = mean_cl_boot, geom = "errorbar", width = 0.1,  position = position_dodge(width = 0.3)) +
  stat_summary(aes(color = Phenotype.x), fun.y = mean, geom = "point", size = 4,  position = position_dodge(width = 0.3)) +
  geom_point(aes(color = Phenotype.x), size = 2, position = position_dodge(width = 0.3)) +
  geom_hline(yintercept=0,linetype=3)+
  ylim(-0.003, 0.002)+
  xlab("Temperature °C")+
  ylab("Pgross nmol/larvae/min")+
  scale_color_manual(values = c("blue", "red"))+
  theme_classic()+
  theme(legend.position = "none")


LEDR <- ggplot(oxygen.bio.devo.wide, aes(x = Temperature.x, y = LEDR )) +
  stat_summary(aes(group = Phenotype.x), fun.y = mean, geom = "path",  position = position_dodge(width = 0.3)) +
  stat_summary(aes(color = Phenotype.x), fun.data = mean_cl_boot, geom = "errorbar", width = 0.1,  position = position_dodge(width = 0.3)) +
  stat_summary(aes(color = Phenotype.x), fun.y = mean, geom = "point", size = 4,  position = position_dodge(width = 0.3)) +
  geom_point(aes(color = Phenotype.x), size = 2, position = position_dodge(width = 0.3)) +
  geom_hline(yintercept=0,linetype=3)+
  ylim(-0.003, 0.002)+
  xlab("Temperature °C")+
  ylab("LEDR nmol/larvae/min")+
  scale_color_manual(values = c("blue", "red"))+
  theme_classic()+
  theme(legend.position = "none")

PR <- ggplot(oxygen.bio.devo.wide, aes(x = Temperature.x, y = PR.ratio )) +
  stat_summary(aes(group = Phenotype.x), fun.y = mean, geom = "path",  position = position_dodge(width = 0.3)) +
  stat_summary(aes(color = Phenotype.x), fun.data = mean_cl_boot, geom = "errorbar", width = 0.1,  position = position_dodge(width = 0.3)) +
  stat_summary(aes(color = Phenotype.x), fun.y = mean, geom = "point", size = 4,  position = position_dodge(width = 0.3)) +
  geom_point(aes(color = Phenotype.x), size = 2, position = position_dodge(width = 0.3)) +
  geom_hline(yintercept=1,linetype=3)+
  ylim(-0.1, 2.6)+
  xlab("Temperature °C")+
  ylab("P:R")+
  scale_color_manual(values = c("blue", "red"))+
  theme_classic()+
  theme(legend.position = c(0.55,0.8))+
  guides(color=guide_legend(title="Devo Temp"))



plots <- grid.arrange(
  Pnet,
  Pgross,
  LEDR,
  PR,
  nrow = 1,
  top = "Pocillopora verrucosa metabolic rates following development at 28°C or 31°C")

ggsave("output/20231212Cohort_DevoTemp_Pverrucosa_Metabolism.pdf", plots , width = 10, height = 4)
ggsave("output/20231212Cohort_DevoTemp_Pverrucosa_Metabolism.jpg", plots , width = 10, height = 4)

```