---
title: "Normalizing Rates"
author: "HM Putnam"
date: "2023-11-26"
output: html_document
---

## Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("devtools" %in% rownames(installed.packages()) == 'FALSE') install.packages('devtools') 
library(devtools)
if ("segmented" %in% rownames(installed.packages()) == 'FALSE') install.packages('segmented') 
if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("LoLinR" %in% rownames(installed.packages()) == 'FALSE') install_github('colin-olito/LoLinR') 
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if ("chron" %in% rownames(installed.packages()) == 'FALSE') install.packages('chron') 
if ("plyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('plyr') 
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr') 
if ("stringr" %in% rownames(installed.packages()) == 'FALSE') install.packages('stringr') 
if ("Rmisc" %in% rownames(installed.packages()) == 'FALSE') install.packages('Rmisc') 
if ("respR" %in% rownames(installed.packages()) == 'FALSE') install.packages('respR') 


#load packages
library("ggplot2")
library("segmented")
library("plotrix")
library("gridExtra")
library("LoLinR")
library("lubridate")
library("chron")
library('plyr')
library('dplyr')
library('stringr')
library('Rmisc')
library('respR')
library('tidyverse')

```

Load Raw extracted Oxygen Flux Rates
```{r, warning=FALSE, message=FALSE}
Photo.Rates <- read.csv(file = "output/pr_sdr_rates_raw.csv") #read file back in so slopes don't have to be generated every time
Photo.Rates = subset(Photo.Rates, select = -c(X) ) #remove empty column

#format  column
Photo.Rates<-Photo.Rates %>%
  mutate(Plate = str_sub(Plate, 1, -1))
Photo.Rates$Plate<-as.integer(Photo.Rates$Plate) #format as # rather than run #, set as integer
```

## Standardize and normalize  

Merge rates files with sample info for normalization, plotting, and stats.  

```{r, warning=FALSE, message=FALSE}
Sample.Info <- read_csv(file = "data/pr_sdr/PR_SDR_Sample_Info.csv")%>%
  select(!Notes)
#make unique ID column
Sample.Info$unique <- paste0(Sample.Info$Date, Sample.Info$Plate, Sample.Info$Chamber.ID)

Photo.Rates$unique <- paste0(Photo.Rates$Date, Photo.Rates$Plate, Photo.Rates$Chamber.ID)

#oxygen<-left_join(Sample.Info, Photo.Rates, by="unique") #add photosynthesis data
oxygen<- left_join(Sample.Info, Photo.Rates, by = "unique", suffix = c("", ".y")) 

colnames(oxygen)[colnames(oxygen) == 'Intercept'] <- 'Photo.Intercept' #rename to specify R
colnames(oxygen)[colnames(oxygen) == 'umol.L.min'] <- 'P.umol.L.min' #rename to specify R

```

Check for Outliers in the raw data
```{r}
oxygen %>% 
  filter(Type=="Blank") %>%
  ggplot(aes(Phenotype, P.umol.L.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

oxygen %>% 
  filter(Type=="Sample") %>%
  ggplot(aes(Phenotype, P.umol.L.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

oxygen %>% 
  ggplot(aes(Type, P.umol.L.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Date, Light, Phenotype), ncol=4)+
  theme_bw()


```

Remove samples that have inaccurate slope extraction for respiration rates (determined by PDF output files and outlier plots and data filled in for "No Sensor" as a test of the code function )  
```{r}
oxygen<-oxygen%>%
  filter(!(unique=="202311174D5")) %>%
  filter(!(unique=="202311175D5")) %>%
  filter(!(unique=="202311195D5")) %>%
  filter(!(unique=="202311176D5")) %>%
  filter(!(unique=="202311196D5"))

oxygen %>% 
  filter(Type=="Blank") %>%
  ggplot(aes(Date, P.umol.L.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

oxygen %>% 
  filter(Type=="Sample") %>%
  ggplot(aes(Date, P.umol.L.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

oxygen %>% 
  ggplot(aes(Type, P.umol.L.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Date, Light, Phenotype), ncol=4)+
  theme_bw()

```

Account for volume to obtain umol per minute.   
```{r, results=TRUE, warning=FALSE, message=FALSE}
#Account for chamber volume to convert from umol L-1 m-1 to umol m-1. This removes per Liter
oxygen$P.umol.min <- oxygen$P.umol.L.min * oxygen$Volume #calculate

```

Check the data from the test of the Number of Larvae needed to get a signal
```{r, results=TRUE, warning=FALSE, message=FALSE}

#filter test data
	
test.number.data <- oxygen %>%
  filter(Date==20231113)

#Subtract blank values. Average blank calculated for each run. Display mean blank 
test_blank_data <- subset(test.number.data, Type == "Blank") #subset to blank data only

test_blank_data %>% 
  ggplot(aes(Light, P.umol.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  theme_bw()

#remove high outlier blanks 
test_blank_data <- test_blank_data %>%
  filter(!P.umol.min>1.0e-04)

#remove low outlier blanks 
test_blank_data <- test_blank_data %>%
  filter(!P.umol.min < (-1.0e-04))

test_blank_data %>% 
  ggplot(aes(Light, P.umol.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  theme_bw()

#display mean blank values
mean.test.blank.data <- test_blank_data %>% group_by(Light) %>%
summarise(mean = mean(P.umol.min, na.rm=TRUE)) #mean P phase blanks

mean.test.blank.data %>% 
  ggplot(aes(Light, mean))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  theme_bw()


test.photo.blnk <- aggregate(P.umol.min ~ Light, data=test_blank_data, mean) #calculate average blank during light for each run
colnames(test.photo.blnk)[colnames(test.photo.blnk) == 'P.umol.min'] <- 'P.Blank.umol.min' #rename to specify blank for R

test.number.data.oxygen <- full_join(test.number.data, test.photo.blnk) #add R blanks to master

test.number.data.oxygen$P.umol.min.corr<-test.number.data.oxygen$P.umol.min-test.number.data.oxygen$P.Blank.umol.min #subtract R blanks

#normalize to larval number
#respiration
test.number.data.oxygen$P.umol.org.min <- test.number.data.oxygen$P.umol.min.corr/test.number.data.oxygen$Org.Number #calculate oxygen per organism 
test.number.data.oxygen$P.nmol.org.min <- test.number.data.oxygen$P.umol.org.min*1000 #calculate nanomoles 

test.number.data.oxygen %>%
  filter(Type=="Sample")%>%
  ggplot(aes(Org.Number, P.umol.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()


#60 larvae generates measureable signals for dark respiration (1) and Photosynthesis (2)

```

Analyze the data by Cohorts without the data for testing by number of larvae 
```{r}
#filter to cohort data
oxygen <- oxygen %>%
  filter(!Date==20231113) %>%
  filter(!Temperature==36)

blank_data <- subset(oxygen, Type == "Blank") #subset to blank data only

blank_data %>% 
  ggplot(aes(Cohort, P.umol.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

#remove high outlier blanks 
blank_data <- blank_data %>%
  filter(!P.umol.min>1.0e-04)

#remove low outlier blanks 
blank_data <- blank_data %>%
  filter(!P.umol.min < (-1.0e-04))

blank_data %>% 
  ggplot(aes(Cohort, P.umol.min))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

#display mean blank values
mean.blank.data <- blank_data %>% group_by(Plate, Light) %>%
summarise(mean = mean(P.umol.min, na.rm=TRUE)) #mean P phase blanks

mean.blank.data %>% 
  ggplot(aes(Plate, mean))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()


photo.blnk <- aggregate(P.umol.min ~ Date*Plate*Light, data=blank_data, mean) #calculate average blank during light for each run
colnames(photo.blnk)[colnames(photo.blnk) == 'P.umol.min'] <- 'P.Blank.umol.min' #rename to specify blank for R

oxygen <- full_join(oxygen, photo.blnk) #add R blanks to master
```

Subtract blank values to generate a "corrected" value for umol O2 min-1.  

```{r, warning=FALSE, message=FALSE}
oxygen$P.umol.min.corr<-oxygen$P.umol.min-oxygen$P.Blank.umol.min #subtract R blanks
```

Normalize to biologically relevant measure. Here, normalize to number of larvae. This can be substituted or changed for larval size/volume as well.  

```{r, warning=FALSE, message=FALSE}
oxygen.bio <- oxygen %>% filter(Type == "Sample") #isolate only biological samples and 
#respiration
oxygen.bio$P.umol.org.min <- oxygen.bio$P.umol.min.corr/oxygen.bio$Org.Number #calculate oxygen per organism 
oxygen.bio$P.nmol.org.min <- oxygen.bio$P.umol.org.min*1000 #calculate nanomoles 

```

Plot values.  
```{r}

oxygen.bio %>% 
  ggplot(aes(Cohort, P.umol.min.corr))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

oxygen.bio$Cohort_Phenotype <- paste0(oxygen.bio$Cohort, "  ", oxygen.bio$Phenotype)


oxygen.bio %>% 
  ggplot(aes(Cohort_Phenotype, P.umol.min.corr))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

```


2-way ANOVA on Cohorts of green fluorescing larvae
```{r}

#filter to the green larvae only
oxygen.bio.green <- oxygen.bio %>%
  filter(!Phenotype == "Red")

oxygen.bio.green %>% 
  ggplot(aes(Cohort, P.umol.min.corr))+
  geom_point()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light))+
  theme_bw()

#subset into photo data and resp data
Photo.bio.green <- oxygen.bio.green %>%
  filter(Light == 1)

mod1 <- lm(P.nmol.org.min ~ as.factor(Cohort), data=Photo.bio.green )
summary(mod1)
hist(mod1$residuals)
plot(mod1$fitted.values, mod1$residuals)

Resp.bio.green <- oxygen.bio.green %>%
  filter(Light == 2)

#remove positive values for respiration
Resp.bio.green <- Resp.bio.green %>%
  filter(P.nmol.org.min < 0 )

mod2 <- lm(P.nmol.org.min ~as.factor(Cohort), data=Resp.bio.green )
summary(mod2)
hist(mod2$residuals)
plot(mod2$fitted.values, mod2$residuals)


plot.oxygen.bio.green <- rbind(Photo.bio.green,Resp.bio.green)

supp.labs <- c("Photosynthesis (Pnet)", "Respiration (LEDR)")
names(supp.labs) <- c("1", "2")

plot.oxygen.bio.green %>% ggplot(., aes(x = Cohort, y = P.nmol.org.min)) +
  geom_point(data = plot.oxygen.bio.green, aes(Cohort, P.nmol.org.min), color="grey") +
  #geom_jitter(width=0.05) +
  stat_summary(fun=mean, shape=1, size=0.5, color="black") +
  stat_summary(fun.data=mean_se, color = 'black', geom = "errorbar", width=0)+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  facet_wrap(vars(Light), labeller = labeller(Light = supp.labs))+
  ylab("Rate (nmol larva-1 min-1)")+
  theme_bw()




```

Save as .csv file.  
```{r, warning=FALSE, message=FALSE}
write.csv(oxygen.bio.green, paste0("output/cohort_pr_calculated_normalized_photo_rates.csv")) #save final file
```


