---
title: "Larval PI Curve Plotting and Analysis" 
author: "AS Huffmyer"
date: '2023'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

QC all plots, check all blanks for every run 

Need to bootstrap for model fit and for parameter estimates 

# Setup  

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 
if ("scales" %in% rownames(installed.packages()) == 'FALSE') install.packages('scales') 
if ("cowplot" %in% rownames(installed.packages()) == 'FALSE') install.packages('cowplot') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("effects" %in% rownames(installed.packages()) == 'FALSE') install.packages('effects') 
if ("emmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('emmeans') 
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp') 
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("padpadpadpad/rTPC")
if ("ggstatsplot" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("IndrajeetPatil/ggstatsplot")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom')

#load packages
library("ggplot2")
library("tidyverse")
library('car')
library('lme4')
library('lmerTest')
library('scales')
library('cowplot')
library('rTPC')
library('ggstatsplot')
library('nls.multstart')
library('broom')
```

# Data visualization and manipulation  

Load data from LoLinR.    
```{r, warning=FALSE, message=FALSE}
pi_data<-read.csv("output/pi_curves/calculated_normalized_photo_rates.csv") #load data
```

Format data. 
```{r, warning=FALSE, message=FALSE}
#remove all rows of wells that did not have samples or blanks
pi_data<-pi_data[!is.na(pi_data$Type),]

#format columns
pi_data$SDR<-as.factor(pi_data$SDR)
pi_data$Plate<-as.factor(pi_data$Plate)
```

Look for outliers in the data.  

```{r}
boxplot(pi_data$P.nmol.org.min)

boxplot(pi_data$P.nmol.org.min~pi_data$PAR)
```

Replace any net negative value with 0, since this indicates no photosynthesis. 
```{r}
pi_data<-pi_data%>%
  mutate(P.nmol.org.min=if_else(P.nmol.org.min<0 & PAR>0, 0, P.nmol.org.min)) 
```

View data. Remove outliers. 
```{r}
boxplot(pi_data$P.nmol.org.min~pi_data$PAR)

pi_data<-pi_data%>%filter(P.nmol.org.min<0.02)

boxplot(pi_data$P.nmol.org.min~pi_data$PAR)
```

Log transform the data.
```{r}
pi_data<-pi_data%>%
  mutate(P.nmol.org.min=log10(P.nmol.org.min+1))

boxplot(pi_data$P.nmol.org.min~pi_data$PAR)
```

Calculate mean temperature values for each run.    

```{r}
pi.temps<-read.csv("output/pi_curves/runs_temp.csv")
pi.temps = subset(pi.temps, select = -c(X) ) #remove empty column

pi_data$Plate<-as.integer(pi_data$Plate)
```


Add temperature data to master data frame.  
```{r}
pi_data<-left_join(pi_data,pi.temps)

#round to 0.1°C 
pi_data<-pi_data%>%
  mutate(Temp.C=round(Temp.C,1))
```

Remove PAR 2 due to sensor malfunction. 
```{r}
pi_data<-pi_data%>%
  filter(!PAR=="4")
```

# Plot photosynthesis over PAR treatments  

Plot data with means   
```{r}
pi_plot1<-pi_data %>%
    group_by(PAR)%>%
    dplyr::summarise(mean=mean(P.nmol.org.min, na.rm=TRUE), sd=sd(P.nmol.org.min, na.rm=TRUE), N=length(P.nmol.org.min), se=sd/sqrt(N))%>%
    
    ggplot(., aes(x = as.factor(PAR), y = mean, group=PAR)) +
    geom_hline(yintercept=0, linetype="dashed", color="black", linewidth=0.75)+
    geom_point(aes(group=PAR), size=6, position = position_dodge(0.4)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se, group=PAR), width=0, linetype="solid", position=position_dodge(0.4), size=0.8, color="black")+
    xlab("Light (PAR)") + 
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pi_plot1

ggsave("figures/pi_curves/pi_means.png", pi_plot1, dpi=300, w=7, h=5, units="in")
```

Plot data with a loess line   
```{r}
pi_plot2<-pi_data %>%
    
    ggplot(., aes(x = as.factor(PAR), y = P.nmol.org.min, group=PAR)) +
    geom_hline(yintercept=0, linetype="dashed", color="black", size=0.75)+
    geom_smooth(aes(group=1), method="loess", alpha=0.4, se=FALSE, colour="black")+
    xlab("Light (PAR)") + 
    ylab(expression(bold(paste("P (nmol ", O[2], " larva"^-1, "min"^-1, ")")))) +
    #ylim(0, 0.05)+
    theme_classic() + 
    theme(
      legend.position="right",
      axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)
      ); pi_plot2
```

# Apply PI curve calculations and models 

## Look for outliers 

View and remove outliers.   
```{r, warning=FALSE, message=FALSE}
#identify outliers by Temperature and Treatment groups
outlier.plot <- ggbetweenstats(pi_data, PAR, P.nmol.org.min, outlier.tagging = TRUE)
outlier.plot

ggsave("figures/pi_curves/outliersbygroup.png", outlier.plot, dpi=300, w=8, h=8, units="in")

#set quantile values
q <- c(0.25, 0.75)

# calculate quantile values by PAR and Treatment groups
Quants <- pi_data %>%
  dplyr::group_by(PAR)%>%
  dplyr::summarise(quant25 = quantile(P.nmol.org.min, probs = q[1], na.rm=TRUE),
            quant75 = quantile(P.nmol.org.min, probs = q[2], na.rm=TRUE),
            IQRbyGroup=IQR(P.nmol.org.min, na.rm=TRUE))


#Calculate Quantile upper and lower ranges 
Quants$upper <-  Quants$quant75+1.5*Quants$IQRbyGroup # Upper Range  
Quants$lower <- Quants$quant25-1.5*Quants$IQRbyGroup # Lower Range

#join outlier cutoffs with rate data
pi_data <- left_join(pi_data, Quants)

#remove outliers from rates
pi_data <- pi_data %>%
  filter(P.nmol.org.min < upper) %>%
  filter(P.nmol.org.min > lower) #%>%
  #filter(rate < 0.125)

# Visualize data following outlier removal
outlier.plot2 <- ggbetweenstats(pi_data, PAR, P.nmol.org.min, outlier.tagging = TRUE)
outlier.plot2

ggsave("figures/pi_curves/removed_outliersbygroup.png", outlier.plot2, dpi=300, w=8, h=8, units="in")
```

## Fitting PI Curves 

Scripts for this section based on E5 PI curve analysis in the time series repository. 

Define data 
```{r}
#specify data
pi_data$PAR <- as.numeric(pi_data$PAR)
pi_data$Pc <- as.numeric(pi_data$P.nmol.org.min)
```

Define PI curve function as a nonlinear Least Squares regression of a quadratic fit, test nls fit

Aquatic Photosynthesis, Falkowski   
Pmax = max photosynthesis (Am from Bayesian script)  
alpha = quantum yeild (AQY from Bayesian script)  
I/E = irradiance (PAR from Bayesian script)  
Rd = dark respiration   
Ik = saturating irradiance 

Next run NLS models.

Using fixed initial values (not using, but keeping for reference):  

```{r}
nest <- nest_legacy
unnest <- unnest_legacy

pi_data$species<-"Pocillopora"

nls_data <- pi_data %>% 
   #mutate(new_Pc=Pc+1)%>%
   group_by(species) %>%
   nest(-species) %>%
   mutate(model1 = map(data, ~ 
                        nls(Pc ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.05,  AQY=0.001, Rd=-0.02)) %>%
                             tidy %>%
                             dplyr::select(term, estimate) %>% 
                             spread(term, estimate))) %>%
  unnest(model1) %>%
  mutate(Ik = Am/AQY) %>%
  mutate(Ic=(Am*Rd)/(AQY*(sqrt(Am^2-Rd^2))))

nls_data

#save 
pi_results<-nls_data

pi_results
```

Using flexible initial values based on input data: not working currently

```{r}
nest <- nest_legacy
unnest <- unnest_legacy

pi_data$species<-"Pocillopora"

#pi_data<-pi_data%>%
  #mutate(new_Pc=Pc+1)

nls_data <- pi_data %>% 
   group_by(species) %>%
   nest(-species) %>%
   mutate(model1 = map(data, ~ 
                         nls(Pc ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=(max(.$Pc)-min(.$Pc)),  AQY=0.001, Rd=-min(.$Pc)), trace=TRUE) %>%
                              tidy %>%
                              dplyr::select(term, estimate) %>% 
                              spread(term, estimate))) %>%
  unnest(model1) %>%
  mutate(Ik = Am/AQY) %>%
  mutate(Ic=(Am*Rd)/(AQY*(sqrt(Am^2-Rd^2))))

#save 
pi_results<-nls_data

pi_results
```

Plot curve over data points.  
```{r}
augmented <- pi_data %>% 
  nest(-species) %>% 
  mutate(
    fit = map(data, ~ nls(Pc ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.5,  AQY=0.001, Rd=.4))),
    augmented = map(fit, augment),
  ) %>% 
  unnest(augmented)

#all together
curve_plot1<-augmented %>%
  group_by(species)%>%
  qplot(PAR, Pc, data = ., geom = 'point') +
  geom_line(aes(y=.fitted), size=2)+
  theme(legend.position="right")+
  theme_classic();curve_plot1

ggsave("figures/pi_curves/nls_curves_all.png", width=6, height=6, dpi=500)
```

Plot PI curve metrics. 

```{r}
metrics_plot1<-pi_results%>%
  pivot_longer(values_to="Value", names_to="Parameter", cols=c(Am:Ic))%>%
  filter(Parameter %in% c("Ik", "Ic"))%>%
  
  ggplot(aes(x=species, y=Value, color=species))+
    facet_wrap(~Parameter)+
    geom_point(aes(), size=4) +
    theme_classic()+
    theme(legend.position="none");metrics_plot1

metrics_plot1a<-pi_results%>%
  pivot_longer(values_to="Value", names_to="Parameter", cols=c(Am:Ic))%>%
  filter(!Parameter=="Ik")%>%
  filter(!Parameter=="Ic")%>%
  
  ggplot(aes(x=species, y=Value, color=species))+
    facet_wrap(~Parameter)+
    geom_point(aes(), size=4) +
    theme_classic();metrics_plot1a

metrics_plot<-plot_grid(metrics_plot1, metrics_plot1a, nrow=1, ncol=2, rel_widths=c(0.8,1))

ggsave("figures/pi_curves/parameters.png", metrics_plot, width=10, height=6, dpi=300)
```




