---
title: "TPC fitting"
authors: Hollie Putnam 
edited by: Hollie Putnam 
date: 20220829
output: html_document
---

# PR data 
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

## install packages if you dont already have them in your library
if (!require("devtools")) install.packages("devtools")
if (!require("furrr")) install.packages("furrr")
if (!require("future")) install.packages("future")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("ggpubr")) install.packages("ggpubr")
if (!require("lubridate")) install.packages("lubridate")
if (!require("cowplot")) install.packages("cowplot")
if (!require("ggstatsplot")) install.packages("ggstatsplot")
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("padpadpadpad/rTPC")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom') 

## load libraries
library('devtools')
library('tidyverse')
library('gridExtra')
library('ggpubr')
library('lubridate')
library('cowplot')
library('ggstatsplot')
library('rTPC')
library('nls.multstart')
library('broom')
library('car')
library('scales')
library('future') ## libraries for parallel processing
library('furrr') ## libraries for parallel processing
library('patchwork')
library('minpack.lm')
```

## Import data
```{r, warning = FALSE}
data <- read.csv("output/pr_curve_extracted_rates.csv")
str(data)
data$Temperature <- as.numeric(data$avg_temp_interval)
str(data)

data <- data %>%
  arrange(colony_id) %>%
  group_by(colony_id) %>%
  mutate(curve_id = group_indices())

data <- data %>%
  filter(Light==560) %>%
  filter(Stage == "Recruit") #%>%
  #filter(Temperature < 39)

data %>% 
ggplot(aes(x=Temperature, y=micromol.cm2.s, color=Species, group=colony_id))+ 
  geom_point()+
  geom_line()+
  ylab("µmol O2 cm^2 s-1")+
  facet_wrap(~ Species*Light,  ncol = 5)+
  theme(legend.position = "none")

```

```{r}
d <- data 

#d$Pc <- replace(d$Pc, d$Pc<0,0)

d <- d[,-c(1,4,6,7,8,9,10,11)]
str(d)

d$curve_id <- sub(".*-", "", d$colony_id)
colnames(d) <- c("Species", "colony_id", "temp","rate","curve_id")

#d$rate <- log10(d$rate+1)

d %>% 
ggplot(aes(x=temp, y=rate, color=Species, group=colony_id))+ 
  geom_point()+
  geom_line()+
  ylab("µmol O2 cm^2 s-1")+
  facet_wrap(~ curve_id,  ncol = 5)+
  theme(legend.position = "none")

d <- d %>%
  filter(Species == "P. meandrina")

 d <- d %>%
   filter(!colony_id == "REC-011") %>% 
   filter(!colony_id == "REC-017") %>%  
   filter(temp < 39)  

d %>% 
ggplot(aes(x=temp, y=rate, color=Species, group=colony_id))+ 
  geom_point()+
  geom_line()+
  ylab("µmol O2 cm^2 s-1")+
  facet_wrap(~ curve_id,  ncol = 5)

d %>% 
ggplot(aes(x=temp, y=rate, color=colony_id, group=colony_id))+ 
  geom_point()+
  geom_line()+
  ylab("µmol O2 cm^2 s-1")
  
  # filter(!colony_id == "REC-009") %>% # need to remove the first data point
  # filter(!colony_id == "REC-023") %>% # need to remove the first data point
  # filter(!colony_id == "REC-031") %>% # need to remove the first data point
  # filter(!colony_id == "REC-030") %>% # need to remove the first data point
  # filter(!colony_id == "REC-035") %>% # need to remove the first data point
  # filter(!colony_id == "REC-037") %>% # need to remove the first data point
  # filter(!colony_id == "REC-039") %>% # need to remove the first data point
  # 
  # filter(!colony_id == "REC-032") %>% # need to remove the first 2 data points
  # 
  # filter(!colony_id == "REC-011") %>% # need to remove the last data point
  # filter(!colony_id == "REC-013") %>% # need to remove the last data point
  # filter(!colony_id == "REC-017") %>% # need to remove the last 4 data points

```


TPC fitting 
Padifeld et al **rTPC and nls.multstart: A new pipeline to fit thermal performance curves in r**  
https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.13585  

Sharpe Schoolfield 1981 model
Schoolfield, R. M., Sharpe, P. J. H., & Magnuson, C. E. (1981). Non-linear regression of biological temperature-dependent rate models based on absolute reaction-rate theory. Journal of theoretical biology, 88(4), 719-731. https://doi.org/10.1016/0022-5193(81)90246-0

#edit nls_multstart to allow for a progress bar
```{r}
# edit nls_multstart to allow for a progress bar
nls_multstart_progress <- function(formula, data = parent.frame(), iter, start_lower, 
                                   start_upper, supp_errors = c("Y", "N"), convergence_count = 100, 
                                   control, modelweights, ...){
  if(!is.null(pb)){
    pb$tick()
  }
  nls_multstart(formula = formula, data = data, iter = iter, start_lower = start_lower, 
                start_upper = start_upper, supp_errors = supp_errors, convergence_count = convergence_count, 
                control = control, modelweights = modelweights, ...)
}
```

# start progress bar and estimate time it will take
```{r}
# start progress bar and estimate time it will take
number_of_models <- 1
number_of_curves <- length(unique(d$curve_id))

# setup progress bar
pb <- progress::progress_bar$new(total = number_of_curves*number_of_models,
                                 clear = FALSE,
                                 format ="[:bar] :percent :elapsedfull")

```

# fit chosen model formulation in rTPC
```{r}
# fit chosen model formulation in rTPC
d_fits <- nest(d, data = c(temp, rate)) %>%
  mutate(sharpeschoolhigh = map(data, ~nls_multstart(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = .x,
                        iter = c(3,3,3,3),
                        start_lower = get_start_vals(.x$temp, .x$rate, model_name = 'sharpeschoolhigh_1981') - 1,
                        start_upper = get_start_vals(.x$temp, .x$rate, model_name = 'sharpeschoolhigh_1981') + 1,
                        lower = get_lower_lims(.x$temp, .x$rate, model_name = 'sharpeschoolhigh_1981'),
                        upper = get_upper_lims(.x$temp, .x$rate, model_name = 'sharpeschoolhigh_1981'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)))

```


```{r}
d_preds <- mutate(d_fits, new_data = map(data, ~tibble(temp = seq(min(.x$temp), max(.x$temp), length.out = 100)))) %>%
  # get rid of original data column
  select(., -data) %>%
  # stack models into a single column, with an id column for model_name
  pivot_longer(., names_to = 'model_name', values_to = 'fit', c(sharpeschoolhigh)) %>%
  # create new list column containing the predictions
  # this uses both fit and new_data list columns
  mutate(preds = map2(fit, new_data, ~augment(.x, newdata = .y))) %>%
  # select only the columns we want to keep
  select(curve_id, Species,model_name, preds) %>%
  # unlist the preds list column
  unnest(preds)

glimpse(d_preds)
```

```{r}
ggplot(d_preds) +
  geom_line(aes(temp, .fitted, col = Species, group=curve_id)) +
  geom_point(aes(temp, rate, col = Species), d) +
  #facet_wrap(~curve_id*Species, scales = 'free_y', ncol = 6) +
  theme_bw() +
  #theme(legend.position = 'none') +
  scale_color_brewer(type = 'qual', palette = 2) +
  labs(x = 'Temperature (ºC)',
       y = 'Photosynthetic rate',
       title = 'sharpeschoolhigh')

ggplot(d_preds) +
  geom_line(aes(temp, .fitted, col = Species, group=curve_id)) +
  geom_point(aes(temp, rate, col = Species), d) +
  facet_wrap(~Species,  ncol = 4) +
  theme_bw() +
  theme(legend.position = 'none') +
  scale_color_brewer(type = 'qual', palette = 2) +
  labs(x = 'Temperature (ºC)',
       y = 'Photosynthetic rate',
       title = 'sharpeschoolhigh')

```

<!-- #calculate TPC metrics by species -->
<!-- ```{r} -->
<!-- d_params <- pivot_longer(d_fits, names_to = 'model_name', values_to = 'fit', c(sharpeschoolhigh)) %>% -->
<!--   mutate(params = map(fit, calc_params)) %>% -->
<!--   select(curve_id, Species, model_name, params) %>% -->
<!--   unnest(params) -->

<!-- glimpse(d_params) -->

<!-- ``` -->


<!-- ```{r} -->
<!-- d_fits <- mutate(d_fits, bootstrap = list(rep(NA, n()))) -->

<!-- # run for loop to bootstrap each refitted model -->
<!-- for(i in 1:nrow(d_fits)){ -->
<!--   temp_data <- d_fits$data[[i]] -->
<!--   temp_fit <- minpack.lm::nlsLM(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28), -->
<!--                         data = temp_data, -->
<!--                         start = coef(d_fits$sharpeschoolhigh[[i]]), -->
<!--                         lower = get_lower_lims(temp_data$temp, temp_data$rate, model_name = 'sharpeschoolhigh_1981')-1, -->
<!--                         upper = get_upper_lims(temp_data$temp, temp_data$rate, model_name = 'sharpeschoolhigh_1981')+1) -->
<!--   boot <- Boot(temp_fit, method = 'residual') -->
<!--   d_fits$bootstrap[[i]] <- boot -->
<!--   rm(list = c('temp_fit', 'temp_data', 'boot')) -->
<!-- } -->

<!-- ``` -->


<!-- ```{r} -->
<!-- # get the raw values of each bootstrap -->
<!-- d_fits <- mutate(d_fits, output_boot = map(bootstrap, function(x) x$t)) -->

<!-- # calculate predictions with a gnarly written function -->
<!-- d_fits <- mutate(d_fits, preds = map2(output_boot, data, function(x, y){ -->
<!--   temp <- as.data.frame(x) %>% -->
<!--     drop_na() %>% -->
<!--     mutate(iter = 1:n()) %>% -->
<!--     group_by_all() %>% -->
<!--     do(data.frame(temp = seq(min(y$temp), max(y$temp), length.out = 100))) %>% -->
<!--     ungroup() %>% -->
<!--     mutate(pred = sharpeschoolhigh_1981(temp, r_tref,e,eh,th, tref=28)) -->
<!--   return(temp) -->
<!-- })) -->

<!-- # select, unnest and calculate 95% CIs of predictions -->
<!-- boot_conf_preds <- select(d_fits, curve_id, preds) %>% -->
<!--   unnest(preds) %>% -->
<!--   group_by(curve_id, temp) %>% -->
<!--   summarise(conf_lower = quantile(pred, 0.025), -->
<!--             conf_upper = quantile(pred, 0.975), -->
<!--             .groups = 'drop') -->

<!-- ggplot() + -->
<!--   geom_line(aes(temp, .fitted), d_preds, col = 'blue') + -->
<!--   geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), boot_conf_preds, fill = 'blue', alpha = 0.3) + -->
<!--   geom_point(aes(temp, rate), d, size = 2) + -->
<!--   theme_bw(base_size = 12) + -->
<!--   labs(x = 'Temperature (ºC)', -->
<!--        y = 'Rate') + -->
<!--   facet_wrap(~curve_id, ncol=2) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- # get tidied parameters using broom::tidy -->
<!-- # get confidence intervals of parameters -->
<!-- d_fits <- mutate(d_fits, params = map(nls_fit, broom::tidy), -->
<!--                  cis = map(bootstrap, function(x){ -->
<!--                    temp <- confint(x, method = 'bca') %>% -->
<!--                      as.data.frame() %>% -->
<!--                      rename(conf_lower = 1, conf_upper = 2) %>% -->
<!--                      rownames_to_column(., var = 'term') -->
<!--                    return(temp) -->
<!--                    })) -->

<!-- # join parameter and confidence intervals in the same dataset  -->
<!-- left_join(select(d_fits, curve_id, growth_temp, flux, params) %>% unnest(params), -->
<!--           select(d_fits, curve_id, growth_temp, flux, cis) %>% unnest(cis)) %>% -->
<!--   ggplot(., aes(curve_id, estimate)) + -->
<!--   geom_point(size = 4) + -->
<!--   geom_linerange(aes(ymin = conf_lower, ymax = conf_upper)) + -->
<!--   theme_bw() + -->
<!--   facet_wrap(~term, scales = 'free') -->
<!-- #> Joining with `by = join_by(curve_id, growth_temp, flux, term)` -->
<!-- ``` -->