---
title: "Photosynthesis TPC fitting"
authors: Hollie Putnam 
edited by: Hollie Putnam 
date: 20220829
output: html_document
---

# Load libraries
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

## install packages if you dont already have them in your library
if (!require("devtools")) install.packages("devtools")
if (!require("furrr")) install.packages("furrr")
if (!require("future")) install.packages("future")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("ggpubr")) install.packages("ggpubr")
if (!require("lubridate")) install.packages("lubridate")
if (!require("cowplot")) install.packages("cowplot")
if (!require("ggstatsplot")) install.packages("ggstatsplot")
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("padpadpadpad/rTPC")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom') 

## load libraries
library('devtools')
library('tidyverse')
library('gridExtra')
library('ggpubr')
library('lubridate')
library('cowplot')
library('ggstatsplot')
library('rTPC')
library('nls.multstart')
library('broom')
library('car')
library('scales')
library('future') ## libraries for parallel processing
library('furrr') ## libraries for parallel processing
library('minpack.lm')
```

## Import data and plot raw data
```{r, warning = FALSE}

#import data
data.0 <- read.csv("output/Nov_pr_curve_extracted_rates.csv")
str(data.0)
data.0$Temperature <- as.numeric(data.0$Temperature)
str(data.0)

#plot raw data grouped by colony for all species
data.0 %>% 
ggplot(aes(x=Temperature, y=micromol.cm2.s, color=Species, group=colony_id))+ 
  geom_point()+
  geom_line()+
  ylab("µmol O2 cm^2 s-1")+
  facet_wrap(~ Species*Light*Stage,  ncol = 6)+
  theme(legend.position = "none")

#filter down to all adults
data <- data.0 %>% 
  filter(Stage=="Adult")

#make a grouping variable
data$group <- paste0(data$colony_id, data$Temperature)

#separate P data
data.light <- data %>%
  filter(Light==560) 

#separate R data
data.dark <- data %>%
  filter(Light==0) 

#joing data and identify temperature of each sample
data2 <- left_join(data.light, data.dark, by = "group") %>%
  select(Species = Species.x, 
         group=group,
         colony_id = colony_id.x, 
         temp = avg_temp_interval.x, 
         temp_dark = avg_temp_interval.y, 
         Stage = Stage.x, 
         Pnet = micromol.cm2.h.x, 
         Rdark = micromol.cm2.h.y)

#calculate Pgross
data2$Pgross <- data2$Pnet- data2$Rdark

#set negative respiration to 0
data2$Rdark <- -data2$Rdark
data2$Rdark <- replace(data2$Rdark, data2$Rdark<0,0)

data2 <- data2 %>%
  filter(!group=="POT-02428") %>%
  filter(!group=="POV-06924")

#plot calculated data for Pgross
data2 %>% 
ggplot(aes(x=temp, y=Pgross, color=Species, group=colony_id))+ 
  geom_point()+
  geom_line()+
  ylab("Pgross µmol O2 cm^2 h-1")+
  facet_wrap(~ Species*colony_id,  ncol = 6)+
  theme(legend.position = "none")

#plot calculated data for Rdark
data2 %>% 
ggplot(aes(x=temp_dark, y=Rdark, color=Species, group=colony_id))+ 
  geom_point()+
  geom_line()+
  ylab("Rdark µmol O2 cm^2 h-1")+
  facet_wrap(~ Species*colony_id,  ncol = 6)+
  theme(legend.position = "none")
```

Photosynthesis
```{r}
#assign dataset to d
d <- data2 

d <- d %>%
  mutate(curve_id = dense_rank(colony_id))  # Assigns a unique rank to each colony_id

#set rate as Pgross
d$rate <- d$Pgross
#range(d$rate)

d <- d %>%
  select(curve_id, Species, temp_dark, rate)  # Replace with the column names you want

#rename colnames
colnames(d) <- c("curve_id", "Species", "temp", "rate")

str(d)

d$curve_id <- as.numeric(d$curve_id)
str(d)


```

#  CURVE FIT 
```{r}

#select data
# d <- d %>% 
#   filter(temp >18) %>% 
#   filter(!curve_id == 3) %>% 
#   filter(!curve_id == 8) %>%
#   filter(!curve_id == 10) %>% 
#   filter(!curve_id == 11) %>% 
#   filter(!curve_id == 12) %>% 
#   filter(!curve_id == 13) %>% 
#   filter(!curve_id == 22) %>% 
#   filter(!curve_id == 23) %>% 
#   filter(!curve_id == 26) %>% 
#   filter(!curve_id == 32)


#plot calculated data for Pgross to check for outliers and data issues
d %>% 
ggplot(aes(x=temp, y=rate, color=curve_id, group=curve_id))+ 
  geom_point()+
  geom_line()+
  ylab("Pgross µmol O2 cm^2 h-1")+
  facet_wrap(~ curve_id,  ncol = 6)+
  theme(legend.position = "right")
```


```{r}
# fit chosen model formulation in rTPC
d_fits <- nest(d, data = c(temp, rate)) %>%
  mutate(sharpeschoolhigh = map(data, ~nls_multstart(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = .x,
                        iter = c(3,3,3,3),
                        start_lower = get_start_vals(.x$temp, .x$rate, model_name = 'sharpeschoolhigh_1981') - 1,
                        start_upper = get_start_vals(.x$temp, .x$rate, model_name = 'sharpeschoolhigh_1981') + 1,
                        lower = get_lower_lims(.x$temp, .x$rate, model_name = 'sharpeschoolhigh_1981'),
                        upper = get_upper_lims(.x$temp, .x$rate, model_name = 'sharpeschoolhigh_1981'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)))

# create new list column of for high resolution data
d_preds <- mutate(d_fits, new_data = map(data, ~tibble(temp = seq(min(.x$temp), max(.x$temp), length.out = 100)))) %>%
  # get rid of original data column
  select(., -data) %>%
  # stack models into a single column, with an id column for model_name
  pivot_longer(., names_to = 'model_name', values_to = 'fit', c(sharpeschoolhigh)) %>%
  # create new list column containing the predictions
  # this uses both fit and new_data list columns
  mutate(preds = map2(fit, new_data, ~augment(.x, newdata = .y))) %>%
  # select only the columns we want to keep
  select(curve_id, Species, model_name, preds) %>%
  # unlist the preds list column
  unnest(preds)

glimpse(d_preds)


# plot
ggplot(d_preds) +
  geom_line(aes(temp, .fitted, col = Species)) +
  geom_point(aes(temp, rate), d) +
  facet_wrap(~curve_id, ncol = 6) +
  theme_bw() +
  theme(legend.position = 'right') +
  scale_color_brewer(type = 'qual', palette = 2) +
  labs(x = 'Temperature (ºC)',
       y = 'Metabolic rate')
```


#calculate all of the TPC parameters
```{r}
d_params <- pivot_longer(d_fits, names_to = 'model_name', values_to = 'fit', c(sharpeschoolhigh)) %>%
  mutate(params = map(fit, calc_params)) %>%
  select(curve_id, Species, model_name, params) %>%
  unnest(params)

str(d_params)

d_params_long <- d_params %>%
  pivot_longer(cols = 4:14,        # Specify the columns to pivot
               names_to = "param", # Name of the new column for parameter names
               values_to = "value") # Name of the new column for parameter values


d_params_long %>%
  filter(!param=="ctmin")  %>%
  filter(!param=="thermal_tolerance")  %>%
ggplot(., aes(x = Species, y = value, color=Species)) +
  geom_point(position = position_jitter(width = 0.2), alpha = 0.7) +  # Data points with jitter for better visibility
  facet_wrap(~param, scales="free_y", ncol = 3) +
  labs(x = "Species",
       y = "value") +
  theme_bw()
```


#extract curve coefficients
```{r}
# get coefs
d_fits <- mutate(d_fits, coefs = map(sharpeschoolhigh, coef))

# fit with nlsLM instead
d_fits <- mutate(d_fits, nls_fit = map2(data, coefs, ~minpack.lm::nlsLM(rate ~ sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                                        data = .x,
                                        start = .y,
                                        lower = get_lower_lims(.x$temp, .x$rate, model_name = 'sharpeschoolhigh_1981')-1,
                                        upper = get_upper_lims(.x$temp, .x$rate, model_name = 'sharpeschoolhigh_1981')+1)))

#view first curve output
d_fits$nls_fit[[1]]

```

Bootstrap each model https://padpadpadpad.github.io/rTPC/articles/bootstrapping_many_curves.html
```{r}

d_fits <- mutate(d_fits, bootstrap = list(rep(NA, n())))

# run for loop to bootstrap each refitted model
for(i in 1:nrow(d_fits)){
  temp_data <- d_fits$data[[i]]
  temp_fit <- minpack.lm::nlsLM(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = temp_data,
                        start = coef(d_fits$sharpeschoolhigh[[i]]),
                        lower = get_lower_lims(temp_data$temp, temp_data$rate, model_name = 'sharpeschoolhigh_1981')-1,
                        upper = get_upper_lims(temp_data$temp, temp_data$rate, model_name = 'sharpeschoolhigh_1981')+1)
  boot <- Boot(temp_fit, method = 'residual')
  d_fits$bootstrap[[i]] <- boot
  rm(list = c('temp_fit', 'temp_data', 'boot'))
}

```

calculate confidence intervals
```{r}
# get the values of each bootstrap
d_fits <- mutate(d_fits, output_boot = map(bootstrap, function(x) x$t))

# calculate predictions 
d_fits <- mutate(d_fits, preds = map2(output_boot, data, function(x, y){
  temp <- as.data.frame(x) %>%
    drop_na() %>%
    mutate(iter = 1:n()) %>%
    group_by_all() %>%
    do(data.frame(temp = seq(min(y$temp), max(y$temp), length.out = 100))) %>%
    ungroup() %>%
    mutate(pred = sharpeschoolhigh_1981(temp, r_tref,e,eh,th, tref=28))
  return(temp)
}))

# select, unnest and calculate 95% CIs of predictions
boot_conf_preds <- select(d_fits, curve_id, preds, Species) %>%
  unnest(preds) %>%
  group_by(curve_id, temp, Species) %>%
  summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975),
            .groups = 'drop')

#Plot fitted curves and CI
 d_preds %>% 
  ggplot() +
  geom_line(aes(temp, .fitted, col = Species)) +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper, fill = Species), boot_conf_preds, alpha = 0.3) +
  geom_point(aes(temp, rate), d, size = 2) +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Rate') +
  facet_wrap(~curve_id, ncol=4)
 
 
 #Plot fitted curves and CI
 d_preds %>% 
  ggplot() +
  geom_line(aes(temp, .fitted, col = Species)) +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper, fill = Species), boot_conf_preds, alpha = 0.3) +
  geom_point(aes(temp, rate), d, size = 2) +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Rate') +
  facet_wrap(~curve_id, ncol=4)
 
```

#extract parameters and bootstrapped CI
```{r}

#Gives error for flat E

# get tidied parameters using broom::tidy
# get confidence intervals of parameters
d_fits <- mutate(d_fits, params = map(nls_fit, broom::tidy),
                 cis = map(bootstrap, function(x){
                   temp <- confint(x, method = 'bca') %>%
                     as.data.frame() %>%
                     rename(conf_lower = 1, conf_upper = 2) %>%
                     rownames_to_column(., var = 'term')
                   return(temp)
                   }))

# join parameter and confidence intervals in the same dataset 
left_join(select(d_fits, curve_id, Species, params) %>% unnest(params),
          select(d_fits, curve_id, Species, params, cis) %>% unnest(cis)) %>%
  ggplot(., aes(Species, estimate, color=Species)) +
  geom_point(size = 4) +
  geom_linerange(aes(ymin = conf_lower, ymax = conf_upper)) +
  theme_bw() +
  facet_wrap(~term, scales = 'free')

```

```{r}

# create empty list column
d_fits <- mutate(d_fits, ci_extra_params = list(rep(NA, n())))


for(i in 1:nrow(d_fits)){
  temp_data <- d_fits$data[[i]]
  temp_fit <- minpack.lm::nlsLM(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = temp_data,
                        start = d_fits$coefs[[i]],
                        lower = get_lower_lims(temp_data$temp, temp_data$rate, model_name = 'sharpeschoolhigh_1981')-1,
                        upper = get_upper_lims(temp_data$temp, temp_data$rate, model_name = 'sharpeschoolhigh_1981')+1)
  boot <- Boot(temp_fit, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(temp_fit)), R = 40, method = 'case') %>%
  confint(., method = 'bca') %>%
  as.data.frame() %>%
  rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param')
  d_fits$ci_extra_params[[i]] <- boot
  rm(list = c('temp_fit', 'temp_data', 'boot'))
}


# calculate extra params for each model and put in long format to begin with
d_fits <- mutate(d_fits, extra_params = map(nls_fit, function(x){calc_params(x) %>% pivot_longer(everything(), names_to =  'param', values_to = 'estimate')}))

dx <- left_join(select(d_fits, curve_id, Species, extra_params) %>% unnest(extra_params),
          select(d_fits, curve_id, Species, ci_extra_params) %>% unnest(ci_extra_params))
  
dx  %>%
  ggplot(., aes(Species, estimate, color=Species)) +
  geom_point(size = 4) +
  geom_linerange(aes(ymin = conf_lower, ymax = conf_upper)) +
  theme_bw() +
  labs(y = 'estimate', x = "curve id") +
  facet_wrap(~param, scales = 'free', ncol=2) +
  labs(title = 'Calculation of confidence intervals for extra parameters')



```

All Species Adults
```{r}
data4 <- d 


```
Pgross

# P. grandis Adult Curve fitting
```{r}
# grandis CURVE FIT
# P. grandis
d.grandis.adult <- data4 %>% 
  filter(Species == "P. grandis")

colnames(d.grandis.adult) <-c("curve_id", "Species", "temp",  "rate")

#fit 
adult.grandis.fit <- nls_multstart(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = d.grandis.adult,
                        iter = c(3,3,3,3),
                        start_lower = get_start_vals(d.grandis.adult$temp, d.grandis.adult$rate, model_name = 'sharpeschoolhigh_1981') - 1,
                        start_upper = get_start_vals(d.grandis.adult$temp, d.grandis.adult$rate, model_name = 'sharpeschoolhigh_1981') + 1,
                        lower = get_lower_lims(d.grandis.adult$temp, d.grandis.adult$rate, model_name = 'sharpeschoolhigh_1981'),
                        upper = get_upper_lims(d.grandis.adult$temp, d.grandis.adult$rate, model_name = 'sharpeschoolhigh_1981'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)

adult.grandis.fit

#generate the predicted data
adult_grandis_new_data <- data.frame(temp = seq(min(d.grandis.adult$temp), max(d.grandis.adult$temp), 0.5))
adult.grandis.preds <- augment(adult.grandis.fit, newdata = adult_grandis_new_data)

#calculate TPC parameters
adult.grandis.TCP.res <- calc_params(adult.grandis.fit) %>%
  mutate_all(round, 2)   # round 

adult.grandis.TCP.res

### Bootstrapping  curve fit    
# refit model using nlsLM
adult.grandis.fit_nlsLM <- minpack.lm::nlsLM(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = d.grandis.adult,
                        start = coef(adult.grandis.fit),
                        lower = get_lower_lims(d.grandis.adult$temp, d.grandis.adult$rate, model_name = 'sharpeschoolhigh_1981')-1,
                        upper = get_upper_lims(d.grandis.adult$temp, d.grandis.adult$rate, model_name = 'sharpeschoolhigh_1981')+1,
                        weights = rep(1, times = nrow(d.grandis.adult)))

# bootstrap using case resampling
adult.grandis.boot1 <- Boot(adult.grandis.fit_nlsLM, method = 'case')

# look at the data
head(adult.grandis.boot1$t)


# create predictions of each bootstrapped model
adult.grandis.boot1_preds <- adult.grandis.boot1$t %>%
  as.data.frame() %>%
  drop_na() %>%
  mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(d.grandis.adult$temp), max(d.grandis.adult$temp), length.out = 100))) %>%
  ungroup() %>%
  mutate(pred = sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 28))

# calculate bootstrapped confidence intervals
adult.grandis.boot1_conf_preds <- group_by(adult.grandis.boot1_preds, temp) %>%
  summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975)) %>%
  ungroup()

# plot bootstrapped CIs
adult.grandis.CI.plot <- ggplot() +
  geom_line(aes(temp, .fitted), adult.grandis.preds, col = '#56B4E9') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), adult.grandis.boot1_conf_preds, fill = '#56B4E9', alpha = 0.3) +
  geom_point(aes(temp, rate), d.grandis.adult, size = 2, alpha = 0.5,col = '#56B4E9') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Rate (µmol O2/cm2/h)')
adult.grandis.CI.plot

```

# P. meandrina Adult Curve fitting
```{r}
# meandrina CURVE FIT
# P. meandrina
d.meandrina.adult <- data4 %>% 
  filter(Species == "P. meandrina")

colnames(d.meandrina.adult) <-c("curve_id", "Species", "temp",  "rate")

#fit 
adult.meandrina.fit <- nls_multstart(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = d.meandrina.adult,
                        iter = c(3,3,3,3),
                        start_lower = get_start_vals(d.meandrina.adult$temp, d.meandrina.adult$rate, model_name = 'sharpeschoolhigh_1981') - 1,
                        start_upper = get_start_vals(d.meandrina.adult$temp, d.meandrina.adult$rate, model_name = 'sharpeschoolhigh_1981') + 1,
                        lower = get_lower_lims(d.meandrina.adult$temp, d.meandrina.adult$rate, model_name = 'sharpeschoolhigh_1981'),
                        upper = get_upper_lims(d.meandrina.adult$temp, d.meandrina.adult$rate, model_name = 'sharpeschoolhigh_1981'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)

adult.meandrina.fit

#generate the predicted data
adult_meandrina_new_data <- data.frame(temp = seq(min(d.meandrina.adult$temp), max(d.meandrina.adult$temp), 0.5))
adult.meandrina.preds <- augment(adult.meandrina.fit, newdata = adult_meandrina_new_data)

#calculate TPC parameters
adult.meandrina.TCP.res <- calc_params(adult.meandrina.fit) %>%
  mutate_all(round, 2)   # round 

adult.meandrina.TCP.res

### Bootstrapping  curve fit    
# refit model using nlsLM
adult.meandrina.fit_nlsLM <- minpack.lm::nlsLM(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = d.meandrina.adult,
                        start = coef(adult.meandrina.fit),
                        lower = get_lower_lims(d.meandrina.adult$temp, d.meandrina.adult$rate, model_name = 'sharpeschoolhigh_1981')-1,
                        upper = get_upper_lims(d.meandrina.adult$temp, d.meandrina.adult$rate, model_name = 'sharpeschoolhigh_1981')+1,
                        weights = rep(1, times = nrow(d.meandrina.adult)))

# bootstrap using case resampling
adult.meandrina.boot1 <- Boot(adult.meandrina.fit_nlsLM, method = 'case')

# look at the data
head(adult.meandrina.boot1$t)


# create predictions of each bootstrapped model
adult.meandrina.boot1_preds <- adult.meandrina.boot1$t %>%
  as.data.frame() %>%
  drop_na() %>%
  mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(d.meandrina.adult$temp), max(d.meandrina.adult$temp), length.out = 100))) %>%
  ungroup() %>%
  mutate(pred = sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 28))

# calculate bootstrapped confidence intervals
adult.meandrina.boot1_conf_preds <- group_by(adult.meandrina.boot1_preds, temp) %>%
  summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975)) %>%
  ungroup()

# plot bootstrapped CIs
adult.meandrina.CI.plot <- ggplot() +
  geom_line(aes(temp, .fitted), adult.meandrina.preds, col = '#0072B2') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), adult.meandrina.boot1_conf_preds, fill = '#0072B2', alpha = 0.3) +
  geom_point(aes(temp, rate), d.meandrina.adult, size = 2, alpha = 0.5,col = '#0072B2') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Rate (µmol O2/cm2/h)')
adult.meandrina.CI.plot

```

# P. tuahiniensis Adult Curve fitting
```{r}
# tuahiniensis CURVE FIT
# P. tuahiniensis
d.tuahiniensis.adult <- data4 %>% 
  filter(Species == "P. tuahiniensis")

colnames(d.tuahiniensis.adult) <-c("curve_id", "Species", "temp",  "rate")

#fit 
adult.tuahiniensis.fit <- nls_multstart(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = d.tuahiniensis.adult,
                        iter = c(3,3,3,3),
                        start_lower = get_start_vals(d.tuahiniensis.adult$temp, d.tuahiniensis.adult$rate, model_name = 'sharpeschoolhigh_1981') - 1,
                        start_upper = get_start_vals(d.tuahiniensis.adult$temp, d.tuahiniensis.adult$rate, model_name = 'sharpeschoolhigh_1981') + 1,
                        lower = get_lower_lims(d.tuahiniensis.adult$temp, d.tuahiniensis.adult$rate, model_name = 'sharpeschoolhigh_1981'),
                        upper = get_upper_lims(d.tuahiniensis.adult$temp, d.tuahiniensis.adult$rate, model_name = 'sharpeschoolhigh_1981'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)

adult.tuahiniensis.fit

#generate the predicted data
adult_tuahiniensis_new_data <- data.frame(temp = seq(min(d.tuahiniensis.adult$temp), max(d.tuahiniensis.adult$temp), 0.5))
adult.tuahiniensis.preds <- augment(adult.tuahiniensis.fit, newdata = adult_tuahiniensis_new_data)

#calculate TPC parameters
adult.tuahiniensis.TCP.res <- calc_params(adult.tuahiniensis.fit) %>%
  mutate_all(round, 2)   # round 

adult.tuahiniensis.TCP.res

### Bootstrapping  curve fit    
# refit model using nlsLM
adult.tuahiniensis.fit_nlsLM <- minpack.lm::nlsLM(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = d.tuahiniensis.adult,
                        start = coef(adult.tuahiniensis.fit),
                        lower = get_lower_lims(d.tuahiniensis.adult$temp, d.tuahiniensis.adult$rate, model_name = 'sharpeschoolhigh_1981')-1,
                        upper = get_upper_lims(d.tuahiniensis.adult$temp, d.tuahiniensis.adult$rate, model_name = 'sharpeschoolhigh_1981')+1,
                        weights = rep(1, times = nrow(d.tuahiniensis.adult)))

# bootstrap using case resampling
adult.tuahiniensis.boot1 <- Boot(adult.tuahiniensis.fit_nlsLM, method = 'case')

# look at the data
head(adult.tuahiniensis.boot1$t)


# create predictions of each bootstrapped model
adult.tuahiniensis.boot1_preds <- adult.tuahiniensis.boot1$t %>%
  as.data.frame() %>%
  drop_na() %>%
  mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(d.tuahiniensis.adult$temp), max(d.tuahiniensis.adult$temp), length.out = 100))) %>%
  ungroup() %>%
  mutate(pred = sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 28))

# calculate bootstrapped confidence intervals
adult.tuahiniensis.boot1_conf_preds <- group_by(adult.tuahiniensis.boot1_preds, temp) %>%
  summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975)) %>%
  ungroup()

# plot bootstrapped CIs
adult.tuahiniensis.CI.plot <- ggplot() +
  geom_line(aes(temp, .fitted), adult.tuahiniensis.preds, col = '#D55E00') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), adult.tuahiniensis.boot1_conf_preds, fill = '#D55E00', alpha = 0.3) +
  geom_point(aes(temp, rate), d.tuahiniensis.adult, size = 2, alpha = 0.5,col = '#D55E00') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Rate (µmol O2/cm2/h)')
adult.tuahiniensis.CI.plot

```


# P. verrucosa Adult Curve fitting
```{r}
# verrucosa CURVE FIT
# P. verrucosa
d.verrucosa.adult <- data4 %>% 
  filter(Species == "P. verrucosa")
d.verrucosa.adult <- d.verrucosa.adult %>% distinct()

colnames(d.verrucosa.adult) <-c("curve_id", "Species", "temp",  "rate")

#fit 
adult.verrucosa.fit <- nls_multstart(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = d.verrucosa.adult,
                        iter = c(3,3,3,3),
                        start_lower = get_start_vals(d.verrucosa.adult$temp, d.verrucosa.adult$rate, model_name = 'sharpeschoolhigh_1981') - 1,
                        start_upper = get_start_vals(d.verrucosa.adult$temp, d.verrucosa.adult$rate, model_name = 'sharpeschoolhigh_1981') + 1,
                        lower = get_lower_lims(d.verrucosa.adult$temp, d.verrucosa.adult$rate, model_name = 'sharpeschoolhigh_1981'),
                        upper = get_upper_lims(d.verrucosa.adult$temp, d.verrucosa.adult$rate, model_name = 'sharpeschoolhigh_1981'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)

adult.verrucosa.fit

#generate the predicted data
adult_verrucosa_new_data <- data.frame(temp = seq(min(d.verrucosa.adult$temp), max(d.verrucosa.adult$temp), 0.5))
adult.verrucosa.preds <- augment(adult.verrucosa.fit, newdata = adult_verrucosa_new_data)

#calculate TPC parameters
adult.verrucosa.TCP.res <- calc_params(adult.verrucosa.fit) %>%
  mutate_all(round, 2)   # round 

adult.verrucosa.TCP.res

### Bootstrapping  curve fit    
# refit model using nlsLM
adult.verrucosa.fit_nlsLM <- minpack.lm::nlsLM(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = d.verrucosa.adult,
                        start = coef(adult.verrucosa.fit),
                        lower = get_lower_lims(d.verrucosa.adult$temp, d.verrucosa.adult$rate, model_name = 'sharpeschoolhigh_1981')-1,
                        upper = get_upper_lims(d.verrucosa.adult$temp, d.verrucosa.adult$rate, model_name = 'sharpeschoolhigh_1981')+1,
                        weights = rep(1, times = nrow(d.verrucosa.adult)))

# bootstrap using case resampling
adult.verrucosa.boot1 <- Boot(adult.verrucosa.fit_nlsLM, method = 'case')

# look at the data
head(adult.verrucosa.boot1$t)


# create predictions of each bootstrapped model
adult.verrucosa.boot1_preds <- adult.verrucosa.boot1$t %>%
  as.data.frame() %>%
  drop_na() %>%
  mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(d.verrucosa.adult$temp), max(d.verrucosa.adult$temp), length.out = 100))) %>%
  ungroup() %>%
  mutate(pred = sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 28))

# calculate bootstrapped confidence intervals
adult.verrucosa.boot1_conf_preds <- group_by(adult.verrucosa.boot1_preds, temp) %>%
  summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975)) %>%
  ungroup()

# plot bootstrapped CIs
adult.verrucosa.CI.plot <- ggplot() +
  geom_line(aes(temp, .fitted), adult.verrucosa.preds, col = '#E69F00') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), adult.verrucosa.boot1_conf_preds, fill = '#E69F00', alpha = 0.3) +
  geom_point(aes(temp, rate), d.verrucosa.adult, size = 2, alpha = 0.5,col = '#E69F00') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Rate (µmol O2/cm2/h)')
adult.verrucosa.CI.plot

```

```{r}
 #set plot colors
# set the colors for plotting
cols <- cbind.data.frame(Species = factor(c("P. tuahiniensis",
                                            "P. meandrina",
                                            "P. verrucosa",
                                            "P. grandis",
                                            "P. cf. effusa",
                                            "P. acuta")),
                         Color = c("#D55E00",
                                   "#0072B2",
                                   "#E69F00",
                                   "#56B4E9",
                                   "#009E73",
                                   "#e63946"))
  
# plot data and model fit
PM.TPC.plot <- ggplot(data=data4, aes(x=temp)) +
   geom_point(aes(temp, rate, color="P. grandis"), d.grandis.adult, col = '#56B4E9',size = 2) +
   geom_point(aes(temp, rate, color="P. meandrina"), d.meandrina.adult, col = '#0072B2', size = 2) +
   geom_point(aes(temp, rate, color="P. tuahiniensis"), d.tuahiniensis.adult, col = '#D55E00',  size = 2) +
   geom_point(aes(temp, rate, color="P. verrucosa"), d.meandrina.adult, col = '#E69F00', size = 2) +
   geom_line(aes(temp, .fitted), adult.grandis.preds, col = '#56B4E9', size=2) +
   geom_line(aes(temp, .fitted), adult.meandrina.preds, col = '#0072B2', size=2) +
   geom_line(aes(temp, .fitted), adult.tuahiniensis.preds, col = '#D55E00', size=2) +
   geom_line(aes(temp, .fitted), adult.verrucosa.preds, col = '#E69F00', size=2) +
   geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), adult.grandis.boot1_conf_preds, fill = "#56B4E9", alpha = 0.3) +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), adult.meandrina.boot1_conf_preds, fill = "#0072B2", alpha = 0.3) +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), adult.tuahiniensis.boot1_conf_preds, fill = "#D55E00", alpha = 0.3) +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), adult.verrucosa.boot1_conf_preds, fill = "#E69F00", alpha = 0.3) +

   xlim(16,41)+
  #ylim(0,0.45)+
   scale_x_continuous(breaks=c(16,22,24,26,28,30,32,34,36,41))+
   theme_bw(base_size = 12) +
   scale_colour_manual(name="Species",values=cols)+
   theme(legend.position = "none",
         panel.border = element_blank(), panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
   labs(x = 'Temperature (ºC)',
        y = expression("Rate"~µmol~O[2] ~cm^{-2}~h^{-1}))

PM.TPC.plot 
```


# Bootstrap CI for all TPC parameters
```{r}
broom::tidy(adult.grandis.fit_nlsLM)
broom::tidy(adult.meandrina.fit_nlsLM)
broom::tidy(adult.tuahiniensis.fit_nlsLM)
broom::tidy(adult.verrucosa.fit_nlsLM)

```


#grandis adults
```{r}
#calculate all the TPC parameters
adult.grandis.extra_params <- calc_params(adult.grandis.fit_nlsLM) %>%
  pivot_longer(everything(), names_to =  'param', values_to = 'estimate')

#calculate CIs for all the TPC parameters
adult.grandis.ci_extra_params <- Boot(adult.grandis.fit_nlsLM, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(adult.grandis.fit_nlsLM)), R = 200, method = 'case') %>%
  confint(., method = 'perc') %>%
  as.data.frame() %>%
  rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'case bootstrap')

#join the parameters and CIs  
adult.grandis.ci_extra_params<- left_join(adult.grandis.ci_extra_params, adult.grandis.extra_params)
adult.grandis.ci_extra_params$Species <- "P. grandis"
```

#meandrina adults
```{r}
#calculate all the TPC parameters
adult.meandrina.extra_params <- calc_params(adult.meandrina.fit_nlsLM) %>%
  pivot_longer(everything(), names_to =  'param', values_to = 'estimate')

#calculate CIs for all the TPC parameters
adult.meandrina.ci_extra_params <- Boot(adult.meandrina.fit_nlsLM, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(adult.meandrina.fit_nlsLM)), R = 200, method = 'case') %>%
  confint(., method = 'perc') %>%
  as.data.frame() %>%
  rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'case bootstrap')

#join the parameters and CIs  
adult.meandrina.ci_extra_params<- left_join(adult.meandrina.ci_extra_params, adult.meandrina.extra_params)
adult.meandrina.ci_extra_params$Species <- "P. meandrina"
```

#tuahiniensis adults
```{r}
#calculate all the TPC parameters
adult.tuahiniensis.extra_params <- calc_params(adult.tuahiniensis.fit_nlsLM) %>%
  pivot_longer(everything(), names_to =  'param', values_to = 'estimate')

#calculate CIs for all the TPC parameters
adult.tuahiniensis.ci_extra_params <- Boot(adult.tuahiniensis.fit_nlsLM, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(adult.tuahiniensis.fit_nlsLM)), R = 200, method = 'case') %>%
  confint(., method = 'perc') %>%
  as.data.frame() %>%
  rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'case bootstrap')

#join the parameters and CIs  
adult.tuahiniensis.ci_extra_params<- left_join(adult.tuahiniensis.ci_extra_params, adult.tuahiniensis.extra_params)
adult.tuahiniensis.ci_extra_params$Species <- "P. tuahiniensis"
```

#verrucosa adults
```{r}
#calculate all the TPC parameters
adult.verrucosa.extra_params <- calc_params(adult.verrucosa.fit_nlsLM) %>%
  pivot_longer(everything(), names_to =  'param', values_to = 'estimate')

#calculate CIs for all the TPC parameters
adult.verrucosa.ci_extra_params <- Boot(adult.verrucosa.fit_nlsLM, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(adult.verrucosa.fit_nlsLM)), R = 200, method = 'case') %>%
  confint(., method = 'perc') %>%
  as.data.frame() %>%
  rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'case bootstrap')

#join the parameters and CIs  
adult.verrucosa.ci_extra_params<- left_join(adult.verrucosa.ci_extra_params, adult.verrucosa.extra_params)
adult.verrucosa.ci_extra_params$Species <- "P. verrucosa"
```

#Join Species estimates and CIs
```{r}
All_params <- rbind(adult.grandis.ci_extra_params, adult.meandrina.ci_extra_params,
                    adult.tuahiniensis.ci_extra_params, adult.verrucosa.ci_extra_params)
All_params <- All_params %>% 
 mutate_if(is.numeric, round, 2)



#Plot all of the estimates
estimate.plots <- ggplot(All_params, aes(Species, estimate, color=Species)) +
  geom_point(size = 2) +
  scale_color_manual(name="Species", values=c("#56B4E9","#0072B2","#D55E00","#E69F00"))+
  geom_linerange(aes(ymin = conf_lower, ymax = conf_upper)) +
  theme_bw() +
  facet_wrap(~param, scales = 'free_y') +
  scale_x_discrete('')

estimate.plots

# #filter to only the most relavent and well characterized parameters
 All_params <- All_params %>% 
   filter(!param=="ctmin") %>%
#   filter(!param=="ctmax") %>%
#   filter(!param=="eh") %>%
#   filter(!param=="rmax") %>%
   filter(!param=="skewness") %>%
#   filter(!param=="topt") %>%
   filter(!param=="thermal_tolerance") #%>%
#   filter(!param=="q10") %>%
#   filter(!param=="e") %>%
#   filter(!param=="breadth")%>%
#   filter(!param=="thermal_safety_margin")
  
#Plot selected  estimates
estimate.plots <- ggplot(All_params, aes(Species, estimate, color=Species)) +
  geom_point(size = 2) +
  scale_color_manual(name="Species", values=c("#56B4E9","#0072B2","#D55E00","#E69F00"))+
  geom_linerange(aes(ymin = conf_lower, ymax = conf_upper)) +
  theme_bw() +
  facet_wrap(~param, scales = 'free_y') +
  scale_x_discrete('')+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

estimate.plots

```

#Plot Curve and Estimate Output
```{r}
#generate a combined figure of TPCs and estimate plots
figure <- ggarrange(PM.TPC.plot , estimate.plots,
                    hjust = c(.01,.01),
                    vjust = c(1.2,.01),
                    labels = c("A) Pgross", "B) BootStrap TPC Estimates"),
                    ncol = 1, nrow = 2,
                    heights=c(1,1))
figure

ggsave("output/Nov_AllSpecies_Pgross_TPC_and_estimates.pdf", figure, dpi=300, w=6, h=8, units="in")
ggsave("output/Nov_AllSpecies_Pgross_TPC_and_estimates.jpg", figure, dpi=300, w=6, h=8, units="in")

```

Rdark

```{r}
#assign dataset to d
d <- data2 

d <- d %>%
  mutate(curve_id = dense_rank(colony_id))  # Assigns a unique rank to each colony_id

#set rate as Pgross
d$rate <- d$Rdark


d <- d %>%
  select(curve_id, Species, temp_dark, rate)  # Replace with the column names you want

#rename colnames
colnames(d) <- c("curve_id", "Species", "temp", "rate")

str(d)

d$curve_id <- as.numeric(d$curve_id)
str(d)

data4 <- d

```

# P. grandis Adult Curve fitting
```{r}
# grandis CURVE FIT
# P. grandis
d.grandis.adult <- data4 %>% 
  filter(Species == "P. grandis")

colnames(d.grandis.adult) <-c("curve_id", "Species", "temp", "rate")

#fit 
adult.grandis.fit <- nls_multstart(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = d.grandis.adult,
                        iter = c(3,3,3,3),
                        start_lower = get_start_vals(d.grandis.adult$temp, d.grandis.adult$rate, model_name = 'sharpeschoolhigh_1981') - 1,
                        start_upper = get_start_vals(d.grandis.adult$temp, d.grandis.adult$rate, model_name = 'sharpeschoolhigh_1981') + 1,
                        lower = get_lower_lims(d.grandis.adult$temp, d.grandis.adult$rate, model_name = 'sharpeschoolhigh_1981'),
                        upper = get_upper_lims(d.grandis.adult$temp, d.grandis.adult$rate, model_name = 'sharpeschoolhigh_1981'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)

adult.grandis.fit

#generate the predicted data
adult_grandis_new_data <- data.frame(temp = seq(min(d.grandis.adult$temp), max(d.grandis.adult$temp), 0.5))
adult.grandis.preds <- augment(adult.grandis.fit, newdata = adult_grandis_new_data)

#calculate TPC parameters
adult.grandis.TCP.res <- calc_params(adult.grandis.fit) %>%
  mutate_all(round, 2)   # round 

adult.grandis.TCP.res

### Bootstrapping  curve fit    
# refit model using nlsLM
adult.grandis.fit_nlsLM <- minpack.lm::nlsLM(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = d.grandis.adult,
                        start = coef(adult.grandis.fit),
                        lower = get_lower_lims(d.grandis.adult$temp, d.grandis.adult$rate, model_name = 'sharpeschoolhigh_1981')-1,
                        upper = get_upper_lims(d.grandis.adult$temp, d.grandis.adult$rate, model_name = 'sharpeschoolhigh_1981')+1,
                        weights = rep(1, times = nrow(d.grandis.adult)))

# bootstrap using case resampling
adult.grandis.boot1 <- Boot(adult.grandis.fit_nlsLM, method = 'case')

# look at the data
head(adult.grandis.boot1$t)


# create predictions of each bootstrapped model
adult.grandis.boot1_preds <- adult.grandis.boot1$t %>%
  as.data.frame() %>%
  drop_na() %>%
  mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(d.grandis.adult$temp), max(d.grandis.adult$temp), length.out = 100))) %>%
  ungroup() %>%
  mutate(pred = sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 28))

# calculate bootstrapped confidence intervals
adult.grandis.boot1_conf_preds <- group_by(adult.grandis.boot1_preds, temp) %>%
  summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975)) %>%
  ungroup()

# plot bootstrapped CIs
adult.grandis.CI.plot <- ggplot() +
  geom_line(aes(temp, .fitted), adult.grandis.preds, col = '#56B4E9') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), adult.grandis.boot1_conf_preds, fill = '#56B4E9', alpha = 0.3) +
  geom_point(aes(temp, rate), d.grandis.adult, size = 2, alpha = 0.5,col = '#56B4E9') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Rate (µmol O2/cm2/h)')
adult.grandis.CI.plot

```

# P. meandrina Adult Curve fitting
```{r}
# meandrina CURVE FIT
# P. meandrina
d.meandrina.adult <- data4 %>% 
  filter(Species == "P. meandrina")

colnames(d.meandrina.adult) <-c("curve_id", "Species", "temp", "rate")

#fit 
adult.meandrina.fit <- nls_multstart(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = d.meandrina.adult,
                        iter = c(3,3,3,3),
                        start_lower = get_start_vals(d.meandrina.adult$temp, d.meandrina.adult$rate, model_name = 'sharpeschoolhigh_1981') - 1,
                        start_upper = get_start_vals(d.meandrina.adult$temp, d.meandrina.adult$rate, model_name = 'sharpeschoolhigh_1981') + 1,
                        lower = get_lower_lims(d.meandrina.adult$temp, d.meandrina.adult$rate, model_name = 'sharpeschoolhigh_1981'),
                        upper = get_upper_lims(d.meandrina.adult$temp, d.meandrina.adult$rate, model_name = 'sharpeschoolhigh_1981'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)

adult.meandrina.fit

#generate the predicted data
adult_meandrina_new_data <- data.frame(temp = seq(min(d.meandrina.adult$temp), max(d.meandrina.adult$temp), 0.5))
adult.meandrina.preds <- augment(adult.meandrina.fit, newdata = adult_meandrina_new_data)

#calculate TPC parameters
adult.meandrina.TCP.res <- calc_params(adult.meandrina.fit) %>%
  mutate_all(round, 2)   # round 

adult.meandrina.TCP.res

### Bootstrapping  curve fit    
# refit model using nlsLM
adult.meandrina.fit_nlsLM <- minpack.lm::nlsLM(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = d.meandrina.adult,
                        start = coef(adult.meandrina.fit),
                        lower = get_lower_lims(d.meandrina.adult$temp, d.meandrina.adult$rate, model_name = 'sharpeschoolhigh_1981')-1,
                        upper = get_upper_lims(d.meandrina.adult$temp, d.meandrina.adult$rate, model_name = 'sharpeschoolhigh_1981')+1,
                        weights = rep(1, times = nrow(d.meandrina.adult)))

# bootstrap using case resampling
adult.meandrina.boot1 <- Boot(adult.meandrina.fit_nlsLM, method = 'case')

# look at the data
head(adult.meandrina.boot1$t)


# create predictions of each bootstrapped model
adult.meandrina.boot1_preds <- adult.meandrina.boot1$t %>%
  as.data.frame() %>%
  drop_na() %>%
  mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(d.meandrina.adult$temp), max(d.meandrina.adult$temp), length.out = 100))) %>%
  ungroup() %>%
  mutate(pred = sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 28))

# calculate bootstrapped confidence intervals
adult.meandrina.boot1_conf_preds <- group_by(adult.meandrina.boot1_preds, temp) %>%
  summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975)) %>%
  ungroup()

# plot bootstrapped CIs
adult.meandrina.CI.plot <- ggplot() +
  geom_line(aes(temp, .fitted), adult.meandrina.preds, col = '#0072B2') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), adult.meandrina.boot1_conf_preds, fill = '#0072B2', alpha = 0.3) +
  geom_point(aes(temp, rate), d.meandrina.adult, size = 2, alpha = 0.5,col = '#0072B2') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Rate (µmol O2/cm2/h)')
adult.meandrina.CI.plot

```

# P. tuahiniensis Adult Curve fitting
```{r}
# tuahiniensis CURVE FIT
# P. tuahiniensis
d.tuahiniensis.adult <- data4 %>% 
  filter(Species == "P. tuahiniensis")

colnames(d.tuahiniensis.adult) <-c("curve_id", "Species", "temp", "rate")

#fit 
adult.tuahiniensis.fit <- nls_multstart(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = d.tuahiniensis.adult,
                        iter = c(3,3,3,3),
                        start_lower = get_start_vals(d.tuahiniensis.adult$temp, d.tuahiniensis.adult$rate, model_name = 'sharpeschoolhigh_1981') - 1,
                        start_upper = get_start_vals(d.tuahiniensis.adult$temp, d.tuahiniensis.adult$rate, model_name = 'sharpeschoolhigh_1981') + 1,
                        lower = get_lower_lims(d.tuahiniensis.adult$temp, d.tuahiniensis.adult$rate, model_name = 'sharpeschoolhigh_1981'),
                        upper = get_upper_lims(d.tuahiniensis.adult$temp, d.tuahiniensis.adult$rate, model_name = 'sharpeschoolhigh_1981'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)

adult.tuahiniensis.fit

#generate the predicted data
adult_tuahiniensis_new_data <- data.frame(temp = seq(min(d.tuahiniensis.adult$temp), max(d.tuahiniensis.adult$temp), 0.5))
adult.tuahiniensis.preds <- augment(adult.tuahiniensis.fit, newdata = adult_tuahiniensis_new_data)

#calculate TPC parameters
adult.tuahiniensis.TCP.res <- calc_params(adult.tuahiniensis.fit) %>%
  mutate_all(round, 2)   # round 

adult.tuahiniensis.TCP.res

### Bootstrapping  curve fit    
# refit model using nlsLM
adult.tuahiniensis.fit_nlsLM <- minpack.lm::nlsLM(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = d.tuahiniensis.adult,
                        start = coef(adult.tuahiniensis.fit),
                        lower = get_lower_lims(d.tuahiniensis.adult$temp, d.tuahiniensis.adult$rate, model_name = 'sharpeschoolhigh_1981')-1,
                        upper = get_upper_lims(d.tuahiniensis.adult$temp, d.tuahiniensis.adult$rate, model_name = 'sharpeschoolhigh_1981')+1,
                        weights = rep(1, times = nrow(d.tuahiniensis.adult)))

# bootstrap using case resampling
adult.tuahiniensis.boot1 <- Boot(adult.tuahiniensis.fit_nlsLM, method = 'case')

# look at the data
head(adult.tuahiniensis.boot1$t)


# create predictions of each bootstrapped model
adult.tuahiniensis.boot1_preds <- adult.tuahiniensis.boot1$t %>%
  as.data.frame() %>%
  drop_na() %>%
  mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(d.tuahiniensis.adult$temp), max(d.tuahiniensis.adult$temp), length.out = 100))) %>%
  ungroup() %>%
  mutate(pred = sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 28))

# calculate bootstrapped confidence intervals
adult.tuahiniensis.boot1_conf_preds <- group_by(adult.tuahiniensis.boot1_preds, temp) %>%
  summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975)) %>%
  ungroup()

# plot bootstrapped CIs
adult.tuahiniensis.CI.plot <- ggplot() +
  geom_line(aes(temp, .fitted), adult.tuahiniensis.preds, col = '#D55E00') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), adult.tuahiniensis.boot1_conf_preds, fill = '#D55E00', alpha = 0.3) +
  geom_point(aes(temp, rate), d.tuahiniensis.adult, size = 2, alpha = 0.5,col = '#D55E00') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Rate (µmol O2/cm2/h)')
adult.tuahiniensis.CI.plot

```


# P. verrucosa Adult Curve fitting
```{r}
# verrucosa CURVE FIT
# P. verrucosa
d.verrucosa.adult <- data4 %>% 
  filter(Species == "P. verrucosa")

colnames(d.verrucosa.adult) <-c("curve_id", "Species", "temp", "rate")

#fit 
adult.verrucosa.fit <- nls_multstart(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = d.verrucosa.adult,
                        iter = c(3,3,3,3),
                        start_lower = get_start_vals(d.verrucosa.adult$temp, d.verrucosa.adult$rate, model_name = 'sharpeschoolhigh_1981') - 1,
                        start_upper = get_start_vals(d.verrucosa.adult$temp, d.verrucosa.adult$rate, model_name = 'sharpeschoolhigh_1981') + 1,
                        lower = get_lower_lims(d.verrucosa.adult$temp, d.verrucosa.adult$rate, model_name = 'sharpeschoolhigh_1981'),
                        upper = get_upper_lims(d.verrucosa.adult$temp, d.verrucosa.adult$rate, model_name = 'sharpeschoolhigh_1981'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)

adult.verrucosa.fit

#generate the predicted data
adult_verrucosa_new_data <- data.frame(temp = seq(min(d.verrucosa.adult$temp), max(d.verrucosa.adult$temp), 0.5))
adult.verrucosa.preds <- augment(adult.verrucosa.fit, newdata = adult_verrucosa_new_data)

#calculate TPC parameters
adult.verrucosa.TCP.res <- calc_params(adult.verrucosa.fit) %>%
  mutate_all(round, 2)   # round 

adult.verrucosa.TCP.res

### Bootstrapping  curve fit    
# refit model using nlsLM
adult.verrucosa.fit_nlsLM <- minpack.lm::nlsLM(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = d.verrucosa.adult,
                        start = coef(adult.verrucosa.fit),
                        lower = get_lower_lims(d.verrucosa.adult$temp, d.verrucosa.adult$rate, model_name = 'sharpeschoolhigh_1981')-1,
                        upper = get_upper_lims(d.verrucosa.adult$temp, d.verrucosa.adult$rate, model_name = 'sharpeschoolhigh_1981')+1,
                        weights = rep(1, times = nrow(d.verrucosa.adult)))

# bootstrap using case resampling
adult.verrucosa.boot1 <- Boot(adult.verrucosa.fit_nlsLM, method = 'case')

# look at the data
head(adult.verrucosa.boot1$t)


# create predictions of each bootstrapped model
adult.verrucosa.boot1_preds <- adult.verrucosa.boot1$t %>%
  as.data.frame() %>%
  drop_na() %>%
  mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(d.verrucosa.adult$temp), max(d.verrucosa.adult$temp), length.out = 100))) %>%
  ungroup() %>%
  mutate(pred = sharpeschoolhigh_1981(temp, r_tref, e, eh, th, tref = 28))

# calculate bootstrapped confidence intervals
adult.verrucosa.boot1_conf_preds <- group_by(adult.verrucosa.boot1_preds, temp) %>%
  summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975)) %>%
  ungroup()

# plot bootstrapped CIs
adult.verrucosa.CI.plot <- ggplot() +
  geom_line(aes(temp, .fitted), adult.verrucosa.preds, col = '#E69F00') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), adult.verrucosa.boot1_conf_preds, fill = '#E69F00', alpha = 0.3) +
  geom_point(aes(temp, rate), d.verrucosa.adult, size = 2, alpha = 0.5,col = '#E69F00') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Rate (µmol O2/cm2/h)')
adult.verrucosa.CI.plot

```

```{r}
 #set plot colors
# set the colors for plotting
cols <- cbind.data.frame(Species = factor(c("P. tuahiniensis",
                                            "P. meandrina",
                                            "P. verrucosa",
                                            "P. grandis",
                                            "P. cf. effusa",
                                            "P. acuta")),
                         Color = c("#D55E00",
                                   "#0072B2",
                                   "#E69F00",
                                   "#56B4E9",
                                   "#009E73",
                                   "#e63946"))
  
# plot data and model fit
PM.TPC.plot <- ggplot(data=data4, aes(x=temp)) +
   geom_point(aes(temp, rate, color="P. grandis"), d.grandis.adult, col = '#56B4E9',size = 2) +
   geom_point(aes(temp, rate, color="P. meandrina"), d.meandrina.adult, col = '#0072B2', size = 2) +
   geom_point(aes(temp, rate, color="P. tuahiniensis"), d.tuahiniensis.adult, col = '#D55E00',  size = 2) +
   geom_point(aes(temp, rate, color="P. verrucosa"), d.meandrina.adult, col = '#E69F00', size = 2) +
   geom_line(aes(temp, .fitted), adult.grandis.preds, col = '#56B4E9', size=2) +
   geom_line(aes(temp, .fitted), adult.meandrina.preds, col = '#0072B2', size=2) +
   geom_line(aes(temp, .fitted), adult.tuahiniensis.preds, col = '#D55E00', size=2) +
   geom_line(aes(temp, .fitted), adult.verrucosa.preds, col = '#E69F00', size=2) +
   geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), adult.grandis.boot1_conf_preds, fill = "#56B4E9", alpha = 0.3) +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), adult.meandrina.boot1_conf_preds, fill = "#0072B2", alpha = 0.3) +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), adult.tuahiniensis.boot1_conf_preds, fill = "#D55E00", alpha = 0.3) +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), adult.verrucosa.boot1_conf_preds, fill = "#E69F00", alpha = 0.3) +

   xlim(16,41)+
  #ylim(0,0.45)+
   scale_x_continuous(breaks=c(16,22,24,26,28,30,32,34,36,41))+
   theme_bw(base_size = 12) +
   scale_colour_manual(name="Species",values=cols)+
   theme(legend.position = "none",
         panel.border = element_blank(), panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
   labs(x = 'Temperature (ºC)',
        y = expression("Rate"~µmol~O[2] ~cm^{-2}~h^{-1}))

PM.TPC.plot 
```


# Bootstrap CI for all TPC parameters
```{r}
broom::tidy(adult.grandis.fit_nlsLM)
broom::tidy(adult.meandrina.fit_nlsLM)
broom::tidy(adult.tuahiniensis.fit_nlsLM)
broom::tidy(adult.verrucosa.fit_nlsLM)

```


#grandis adults
```{r}
#calculate all the TPC parameters
adult.grandis.extra_params <- calc_params(adult.grandis.fit_nlsLM) %>%
  pivot_longer(everything(), names_to =  'param', values_to = 'estimate')

#calculate CIs for all the TPC parameters
adult.grandis.ci_extra_params <- Boot(adult.grandis.fit_nlsLM, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(adult.grandis.fit_nlsLM)), R = 200, method = 'case') %>%
  confint(., method = 'perc') %>%
  as.data.frame() %>%
  rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'case bootstrap')

#join the parameters and CIs  
adult.grandis.ci_extra_params<- left_join(adult.grandis.ci_extra_params, adult.grandis.extra_params)
adult.grandis.ci_extra_params$Species <- "P. grandis"
```

#meandrina adults
```{r}
#calculate all the TPC parameters
adult.meandrina.extra_params <- calc_params(adult.meandrina.fit_nlsLM) %>%
  pivot_longer(everything(), names_to =  'param', values_to = 'estimate')

#calculate CIs for all the TPC parameters
adult.meandrina.ci_extra_params <- Boot(adult.meandrina.fit_nlsLM, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(adult.meandrina.fit_nlsLM)), R = 200, method = 'case') %>%
  confint(., method = 'perc') %>%
  as.data.frame() %>%
  rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'case bootstrap')

#join the parameters and CIs  
adult.meandrina.ci_extra_params<- left_join(adult.meandrina.ci_extra_params, adult.meandrina.extra_params)
adult.meandrina.ci_extra_params$Species <- "P. meandrina"
```

#tuahiniensis adults
```{r}
#calculate all the TPC parameters
adult.tuahiniensis.extra_params <- calc_params(adult.tuahiniensis.fit_nlsLM) %>%
  pivot_longer(everything(), names_to =  'param', values_to = 'estimate')

#calculate CIs for all the TPC parameters
adult.tuahiniensis.ci_extra_params <- Boot(adult.tuahiniensis.fit_nlsLM, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(adult.tuahiniensis.fit_nlsLM)), R = 200, method = 'case') %>%
  confint(., method = 'perc') %>%
  as.data.frame() %>%
  rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'case bootstrap')

#join the parameters and CIs  
adult.tuahiniensis.ci_extra_params<- left_join(adult.tuahiniensis.ci_extra_params, adult.tuahiniensis.extra_params)
adult.tuahiniensis.ci_extra_params$Species <- "P. tuahiniensis"
```

#verrucosa adults
```{r}
#calculate all the TPC parameters
adult.verrucosa.extra_params <- calc_params(adult.verrucosa.fit_nlsLM) %>%
  pivot_longer(everything(), names_to =  'param', values_to = 'estimate')

#calculate CIs for all the TPC parameters
adult.verrucosa.ci_extra_params <- Boot(adult.verrucosa.fit_nlsLM, f = function(x){unlist(calc_params(x))}, labels = names(calc_params(adult.verrucosa.fit_nlsLM)), R = 200, method = 'case') %>%
  confint(., method = 'perc') %>%
  as.data.frame() %>%
  rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'case bootstrap')

#join the parameters and CIs  
adult.verrucosa.ci_extra_params<- left_join(adult.verrucosa.ci_extra_params, adult.verrucosa.extra_params)
adult.verrucosa.ci_extra_params$Species <- "P. verrucosa"
```

#Join Species estimates and CIs
```{r}
All_params <- rbind(adult.grandis.ci_extra_params, adult.meandrina.ci_extra_params,
                    adult.tuahiniensis.ci_extra_params, adult.verrucosa.ci_extra_params)
All_params <- All_params %>% 
 mutate_if(is.numeric, round, 2)



#Plot all of the estimates
estimate.plots <- ggplot(All_params, aes(Species, estimate, color=Species)) +
  geom_point(size = 2) +
  scale_color_manual(name="Species", values=c("#56B4E9","#0072B2","#D55E00","#E69F00"))+
  geom_linerange(aes(ymin = conf_lower, ymax = conf_upper)) +
  theme_bw() +
  facet_wrap(~param, scales = 'free_y') +
  scale_x_discrete('')

estimate.plots

# #filter to only the most relavent and well characterized parameters
 All_params <- All_params %>% 
   filter(!param=="ctmin") %>%
#   filter(!param=="ctmax") %>%
#   filter(!param=="eh") %>%
#   filter(!param=="rmax") %>%
   filter(!param=="skewness") %>%
#   filter(!param=="topt") %>%
   filter(!param=="thermal_tolerance") #%>%
#   filter(!param=="q10") %>%
#   filter(!param=="e") %>%
#   filter(!param=="breadth")%>%
#   filter(!param=="thermal_safety_margin")
  
#Plot selected  estimates
estimate.plots <- ggplot(All_params, aes(Species, estimate, color=Species)) +
  geom_point(size = 2) +
  scale_color_manual(name="Species", values=c("#56B4E9","#0072B2","#D55E00","#E69F00"))+
  geom_linerange(aes(ymin = conf_lower, ymax = conf_upper)) +
  theme_bw() +
  facet_wrap(~param, scales = 'free_y') +
  scale_x_discrete('')+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

estimate.plots

```

#Plot Curve and Estimate Output
```{r}
#generate a combined figure of TPCs and estimate plots
figure <- ggarrange(PM.TPC.plot , estimate.plots,
                    hjust = c(.01,.01),
                    vjust = c(1.2,.01),
                    labels = c("C) Rdark", "D) BootStrap TPC Estimates"),
                    ncol = 1, nrow = 2,
                    heights=c(1,1))
figure

ggsave("output/AllSpecies_Rdark_TPC_and_estimates.pdf", figure, dpi=300, w=6, h=8, units="in")
ggsave("output/AllSpecies_Rdark_TPC_and_estimates.jpg", figure, dpi=300, w=6, h=8, units="in")

```