---
title: "Photosynthesis TPC fitting"
authors: Hollie Putnam 
edited by: Hollie Putnam 
date: 20220829
output: html_document
---

# PR data from heatwave timepoint 1 corals
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

## install packages if you dont already have them in your library
if (!require("devtools")) install.packages("devtools")
if (!require("furrr")) install.packages("furrr")
if (!require("future")) install.packages("future")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("ggpubr")) install.packages("ggpubr")
if (!require("lubridate")) install.packages("lubridate")
if (!require("cowplot")) install.packages("cowplot")
if (!require("ggstatsplot")) install.packages("ggstatsplot")
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("padpadpadpad/rTPC")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom') 

## load libraries
library('devtools')
library('tidyverse')
library('gridExtra')
library('ggpubr')
library('lubridate')
library('cowplot')
library('ggstatsplot')
library('rTPC')
library('nls.multstart')
library('broom')
library('car')
library('scales')
library('future') ## libraries for parallel processing
library('furrr') ## libraries for parallel processing
```

## Import data
```{r, warning = FALSE}
data <- read.csv("output/pr_curve_extracted_rates.csv")
str(data)
data$Temperature <- as.numeric(data$Temperature)
str(data)

data %>% 
ggplot(aes(x=Temperature, y=micromol.cm2.s, color=Species, group=colony_id.x))+ 
  geom_point()+
  geom_line()+
  ylab("µmol O2 cm^2 s-1")+
  facet_wrap(~ Species*Light,  ncol = 5)+
  theme(legend.position = "none")

```

```{r}
d <- data %>%
  filter(Light==560) %>%
  filter(!Species == "Recruit") %>%
  #filter(!colony_id.x == "POT-030")%>%
  filter(!colony_id.x == "POT-090")%>%
  #filter(!colony_id.x == "POV-004")%>%
  filter(!colony_id.x == "POV-069")

#d$micromol.cm2.s <- d$micromol.cm2.s
d$micromol.cm2.s <- replace(d$micromol.cm2.s, d$micromol.cm2.s<0,0)

d <- d[,-c(1,3,6,7,9)]
str(d)

colnames(d) <- c("Species", "curve_id", "temp","rate")
d$curve_id <- as.numeric(d$curve_id)
str(d)

```


TPC fitting 
Padifeld et al **rTPC and nls.multstart: A new pipeline to fit thermal performance curves in r**  
https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.13585  

Sharpe Schoolfield 1981 model
Schoolfield, R. M., Sharpe, P. J. H., & Magnuson, C. E. (1981). Non-linear regression of biological temperature-dependent rate models based on absolute reaction-rate theory. Journal of theoretical biology, 88(4), 719-731. https://doi.org/10.1016/0022-5193(81)90246-0

#edit nls_multstart to allow for a progress bar
```{r}
# edit nls_multstart to allow for a progress bar
nls_multstart_progress <- function(formula, data = parent.frame(), iter, start_lower, 
                                   start_upper, supp_errors = c("Y", "N"), convergence_count = 100, 
                                   control, modelweights, ...){
  if(!is.null(pb)){
    pb$tick()
  }
  nls_multstart(formula = formula, data = data, iter = iter, start_lower = start_lower, 
                start_upper = start_upper, supp_errors = supp_errors, convergence_count = convergence_count, 
                control = control, modelweights = modelweights, ...)
}
```

# start progress bar and estimate time it will take
```{r}
# start progress bar and estimate time it will take
number_of_models <- 1
number_of_curves <- length(unique(d$curve_id))

# setup progress bar
pb <- progress::progress_bar$new(total = number_of_curves*number_of_models,
                                 clear = FALSE,
                                 format ="[:bar] :percent :elapsedfull")

```

# fit chosen model formulation in rTPC
```{r}
# fit chosen model formulation in rTPC
d_fits <- nest(d, data = c(temp, rate)) %>%
  mutate(sharpeschoolhigh = map(data, ~nls_multstart_progress(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 28),
                        data = .x,
                        iter = c(3,3,3,3),
                        start_lower = get_start_vals(.x$temp, .x$rate, model_name = 'sharpeschoolhigh_1981') - 10,
                        start_upper = get_start_vals(.x$temp, .x$rate, model_name = 'sharpeschoolhigh_1981') + 10,
                        lower = get_lower_lims(.x$temp, .x$rate, model_name = 'sharpeschoolhigh_1981'),
                        upper = get_upper_lims(.x$temp, .x$rate, model_name = 'sharpeschoolhigh_1981'),
                        supp_errors = 'Y',
                        convergence_count = FALSE)))

```


```{r}
d_preds <- mutate(d_fits, new_data = map(data, ~tibble(temp = seq(min(.x$temp), max(.x$temp), length.out = 100)))) %>%
  # get rid of original data column
  select(., -data) %>%
  # stack models into a single column, with an id column for model_name
  pivot_longer(., names_to = 'model_name', values_to = 'fit', c(sharpeschoolhigh)) %>%
  # create new list column containing the predictions
  # this uses both fit and new_data list columns
  mutate(preds = map2(fit, new_data, ~augment(.x, newdata = .y))) %>%
  # select only the columns we want to keep
  select(curve_id, Species,model_name, preds) %>%
  # unlist the preds list column
  unnest(preds)

glimpse(d_preds)
```

```{r}
ggplot(d_preds) +
  geom_line(aes(temp, .fitted, col = Species, group=curve_id)) +
  geom_point(aes(temp, rate, col = Species), d) +
  #facet_wrap(~curve_id*Species, scales = 'free_y', ncol = 6) +
  theme_bw() +
  theme(legend.position = 'none') +
  scale_color_brewer(type = 'qual', palette = 2) +
  labs(x = 'Temperature (ºC)',
       y = 'Photosynthetic rate',
       title = 'sharpeschoolhigh')

ggplot(d_preds) +
  geom_line(aes(temp, .fitted, col = Species, group=curve_id)) +
  geom_point(aes(temp, rate, col = Species), d) +
  facet_wrap(~curve_id*Species, scales = 'free_y', ncol = 2) +
  theme_bw() +
  theme(legend.position = 'none') +
  scale_color_brewer(type = 'qual', palette = 2) +
  labs(x = 'Temperature (ºC)',
       y = 'Photosynthetic rate',
       title = 'sharpeschoolhigh')

```